### 2023-03-16 Manuscript-figures.Rmd
Updating scripts used in manuscript

---
title: "FireSim2019_ManuscriptFigures"
author: "Dana Johnson"
date: "9/9/2021"
output: html_document
---
```{r setup, include=FALSE}

#library(knitr)
library(ggplot2)
#library(plyr)
library(dplyr)
library(tidyr)
#library(car) # For Levene's Test
library(phyloseq)
#library(scatterplot3d)
#library(ggtree)
library(cowplot) # add-on to ggplot to assist in publication-quality figs
#library(qwraps2)
library(vegan)
library(tidyverse)
#library(psych) # scatterplot matrix (pairs, or pairs.panels)
library(ggtern) # for ternary plots
#library(devtools)
#library(DivNet)
#library(breakaway) # for richness estimates

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(out.width = '100%')

```

Defaults = theme_bw(), black = O, "organic"; grey40 - M, "mineral"; shapes = circle, triangle

# Metadata
```{r, include=FALSE}
df.meta <- read.csv('../../data/all-sample-metadata.csv')

df.meta <- subset(df.meta, core.id != '19UW-WB-11-02' &
                         Full.id != '19UW-WB-06-08-A-SI-duplicate' &
                         Full.id != '19UW-WB-07-02-O-SI-duplicate' &
                         Full.id != '19UW-WB-11-03-O-SI-duplicate' &
                         Full.id != '19UW-WB-19-07-A-SI-duplicate' &
                         Full.id != '19UW-WB-08-10-O-SI')

# Assign degree hours by horizon:
for (i in 1:nrow(df.meta)) {
  if (df.meta$PRE.O.hor.thickness.cm[i] == 10) {
    df.meta$hrzn.degree.hrs[i] = (df.meta$Low.degree.C.hours[i] + df.meta$Mid.degree.C.hours[i])/2
  } else if (df.meta$horizon[i] == 'A') {
    df.meta$hrzn.degree.hrs[i] = df.meta$Low.degree.C.hours[i]
  } else {
    df.meta$hrzn.degree.hrs[i] = df.meta$Mid.degree.C.hours[i]
  }
}

# Create labels and palettes for plots
HorizonLabels = c('Organic','Mineral')
names(HorizonLabels) = c('O','A')

DNALabels   = c('RNA','DNA')
names(DNALabels) = c('cDNA','gDNA')

ComparisonLabels = c('dry vs. unburned','wet vs. unburned')
names(ComparisonLabels) = c("DvC",'WvC')

# Edit names of Leading Species
VegLabels_Condensed = c("Picea spp.","Pinus banksiana","Populus \n tremuloides")
names(VegLabels_Condensed) <- c("Picea","Pinus_banksiana","Populus_tremuloides")

VegLabels_Full = c("Picea glauca","Picea mariana","Pinus banksiana","Pinus banksiana \n overstory","Populus \n tremuloides")
names(VegLabels_Full) <- c("Picea_glauca","Picea_mariana","Pinus_banksiana","Pinus_banksiana overstory piceglauc understory","Populus_tremula")

ThermoLabels = c('Organic-mineral \ninterface','Core base')
names(ThermoLabels) <- c('Mid','Low')

BurnLabels = c('Dry soil burn','Wet soil burn', 'Control')
names(BurnLabels) <- c('dry','wet', 'control')

# Edit names of Site
SiteLabels <- c("Site 01","Site 02","Site 03","Site 04","Site 05","Site 06",
               "Site 07","Site 08","Site 09","Site 10","Site 11","Site 12",
               "Site 13","Site 14","Site 15","Site 16","Site 17","Site 18",
               "Site 19")
names(SiteLabels) <- c("1","2","3","4","5","6","7","8","9","10","11",
                      "12","13","14","15","16","17","18","19")

BurnSeverityLabels = c('Wet burn','Dry burn')
names(BurnSeverityLabels) <- c("Low_Sev",'High_Sev')

IncubLabels = c('24 hours post-burn','5 weeks post-burn', '6 months post-burn', '6 months post-burn: \nAutoclaved','6 months post-burn: \nNot autoclaved')
names(IncubLabels) = c('pb','SI', 'LI', 'LIwA','LIn')

CoefsLabel = c('Slow pool (M2)', 'Fast pool (M1)', 'Slow pool decay rate (k2)','Fast pool decay rate (k1)')
names(CoefsLabel) = c('M2','M1', 'k2','k1')

# Reorder burn treatment levels
df.meta$horizon =factor(df.meta$horizon, levels = c('O','A'))
df.meta$incub.trtmt = factor(df.meta$incub.trtmt, levels = c('pb','SI','LIwA','LIn'))
df.meta$burn.trtmt = factor(df.meta$burn.trtmt, levels = c('control','wet','dry'))
```

# Load phyloseq objects
```{r, message = FALSE}
#df.pb.gDNA.OTU <- read.csv('../data/sequence-data/LibCombined/phyloseq-objects/df.pb.gDNA.norm.csv')
#df.pb.cDNA.OTU <- read.csv('../data/sequence-data/LibCombined/phyloseq-objects/df.pb.cDNA.norm.csv')
#df.SI.OTU <- read.csv('../data/sequence-data/LibCombined/phyloseq-objects/df.SI.norm.csv')
#df.LI.norm <- read.csv('../data/sequence-data/LibCombined/phyloseq-objects/df.LI.norm.csv')

# Starting phyloseq:
### Normalized + tree
ps.norm.full <- readRDS('../../data/sequence-data/LibCombined/phyloseq-objects/ps.norm.full')
ps.norm.full <- prune_taxa(taxa_sums(ps.norm.full)>0, ps.norm.full)

ps.norm.full <- prune_samples(sample_data(ps.norm.full)$Full.id != '19UW-WB-06-08-A-SI-duplicate' &
                    sample_data(ps.norm.full)$Full.id != '19UW-WB-07-02-O-SI-duplicate' &
                    sample_data(ps.norm.full)$Full.id != '19UW-WB-11-03-O-SI-duplicate'&
                    sample_data(ps.norm.full)$Full.id != '19UW-WB-19-07-A-SI-duplicate' &
                    sample_data(ps.norm.full)$Full.id != '19UW-WB-08-10-O-SI', ps.norm.full)

ps.raw.full <- readRDS('../../data/sequence-data/LibCombined/phyloseq-objects/ps.raw.full')

ps.raw.full <- prune_samples(sample_data(ps.raw.full)$Full.id != '19UW-WB-06-08-A-SI-duplicate' &
                    sample_data(ps.raw.full)$Full.id != '19UW-WB-07-02-O-SI-duplicate' &
                    sample_data(ps.raw.full)$Full.id != '19UW-WB-11-03-O-SI-duplicate'&
                    sample_data(ps.raw.full)$Full.id != '19UW-WB-19-07-A-SI-duplicate' &
                    sample_data(ps.raw.full)$Full.id != '19UW-WB-08-10-O-SI', ps.raw.full)
ps.raw.full <- prune_taxa(taxa_sums(ps.raw.full)>0, ps.raw.full)

# Reorder vegetation:
sample_data(ps.norm.full)$Veg.type = factor(sample_data(ps.norm.full)$Veg.type, levels = c('Picea','Populus_tremuloides','Pinus_banksiana'))
sample_data(ps.norm.full)$horizon = factor(sample_data(ps.norm.full)$horizon, levels = c('O','A'))
sample_data(ps.norm.full)$incub.trtmt = factor(sample_data(ps.norm.full)$incub.trtmt, levels = c('pb', 'SI','LIwA','LIn'))

#Pull out just the gDNA samples:
ps.norm.gDNA <- prune_samples(sample_data(ps.norm.full)$DNA.type == 'gDNA', ps.norm.full)
ps.raw.gDNA <- prune_samples(sample_data(ps.raw.full)$DNA.type == 'gDNA', ps.raw.full)

ps.norm.cDNA <- prune_samples(sample_data(ps.norm.full)$DNA.type == 'cDNA', ps.norm.full)
ps.raw.cDNA <- prune_samples(sample_data(ps.raw.full)$DNA.type == 'cDNA', ps.raw.full)

#ps.raw.pb <- prune_samples(sample_data(ps.raw.full)$incub.trtmt == 'pb', ps.raw.full)
#ps.norm.pb <- prune_samples(sample_data(ps.norm.full)$incub.trtmt == 'pb', ps.norm.full)

# ps.norm.pb
#ps.norm.pb.gDNA <- prune_samples(sample_data(ps.norm.gDNA)$incub.trtmt == 'pb',ps.norm.gDNA)
#ps.norm.pb.cDNA <- prune_samples(sample_data(ps.norm.cDNA)$incub.trtmt == 'pb',ps.norm.cDNA)
# ps.norm.SI
#ps.norm.SI <- prune_samples(sample_data(ps.norm.gDNA)$incub.trtmt == 'SI',ps.norm.gDNA)
# ps.norm.LIwA
#ps.norm.LIwA <- prune_samples(sample_data(ps.norm.gDNA)$incub.trtmt == 'LIwA',ps.norm.gDNA)
#ps.norm.LIn <- prune_samples(sample_data(ps.norm.gDNA)$incub.trtmt == 'LIn',ps.norm.gDNA)
#ps.norm.LI <- prune_samples(sample_data(ps.norm.gDNA)$incub.trtmt %in% c('LIn', 'LIwA'),ps.norm.gDNA)
#ps.raw.LI <- prune_samples(sample_data(ps.raw.gDNA)$incub.trtmt %in% c('LIn', 'LIwA'),ps.raw.gDNA)

#ps.norm.SI<- prune_taxa(taxa_sums(ps.norm.SI)>0, ps.norm.SI)
```

# Figure 2 (burn temperature, pH and C:N change, respiration coefficients):
```{r}
# Temperatures ----
df.Temp <-read.csv('../../data/burn-thermocouple-temp-logs/Temp-all-burns-clean.csv')
df.Temp$Thermo_position <- factor(df.Temp$Thermo_position, levels = c('Mid','Low'))

pTemp = ggplot(df.Temp, aes(x = time_hr, 
                          y = Temp)) + 
  geom_point(aes(color = burn.trtmt), size=1) +
  theme_bw()+
  facet_wrap(~Thermo_position,scales = 'free_y', ncol=1,
             labeller = labeller(Thermo_position = ThermoLabels,
                                 burn.trtmt = BurnLabels)) +
  labs(x = "Time (h)",
       y = expression(paste("Thermocouple temperature ( ",degree ~ C, ")")),
       color = "Burn treatment") + 
  #scale_color_manual(values = c("black","grey70"),
   #                  labels = c('O horizon/mineral \n interface','9 cm'),
    #                 breaks = c('wet','dry')) + 
  scale_color_manual(values = c('orange','red'),
                     limits = c('wet','dry'),
                     labels = c('Moist soil burn','Dry soil burn')) +
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        axis.text = element_text(size=14),
        strip.text.y = element_text(size=14), 
        strip.text.x = element_text(size=14), 
        legend.text = element_text(size=12),
        legend.title = element_text(size=12), 
        legend.position = 'right',
       strip.background = element_rect(colour="black", fill="grey92")) +
  geom_hline(yintercept=100)
pTemp

# pH and C:N change ----
pChange_pH <- ggplot(subset(df.meta, incub.trtmt == 'SI')) + 
  geom_boxplot(aes(x=horizon, y = pH, fill = burn.trtmt), varwidth = FALSE) + 
    theme_bw()+
  labs(x= 'Soil horizon',
     #  y = 'Decay rate constant for fast C pool \nlog(k1)',
       y = 'Soil pH',
       fill = 'Burn treatment') + 
  scale_fill_manual(values = c('grey50','orange','red'),
                     limits = c('control','wet','dry'),
                     labels = c('Unburned\ncontrol','Moist soil\nburn','Dry soil\nburn')) +
  scale_x_discrete(limits = c('O','A'),
                     labels = c('Organic','Mineral')) +
  #scale_fill_manual(values = c("deepskyblue3","saddlebrown"),
  # scale_fill_manual(values = c("slateblue4","palegreen3"),
  # #scale_fill_manual(values = c("darkgreen","saddlebrown"),
  #                    limits = c('O','A'),
  #                    labels = c('Organic','Mineral')) +
  # scale_x_discrete(limits = c('control','wet','dry'),
  #                    labels = c('Unburned\ncontrol','Wet soil\nburn','Dry soil\nburn')) +
  theme(axis.text.y = element_text(size = 14),
        legend.title = element_text(size=14),
        axis.text.x = element_text(size = 14),
        axis.title.x = element_text(size = 14, margin = margin(t=10, b=10)),
        axis.title.y = element_text(size = 14, margin = margin(r=10, l=10)),
        strip.text = element_text(size=14),
        legend.text = element_text(size=14),
        legend.position = '',
       # legend.background = element_rect(fill = "darkgray"),
        legend.box = "vertical")

pChange_CN  <- ggplot(subset(df.meta, incub.trtmt == 'SI' &site !=10)) + 
  geom_boxplot(aes(x=horizon, y = C.N.ratio, fill = burn.trtmt), varwidth = FALSE) + 
    theme_bw()+
  labs(x= 'Soil horizon',
     #  y = 'Decay rate constant for fast C pool \nlog(k1)',
       y = 'Soil C:N',
       fill = 'Burn treatment') + 
  scale_fill_manual(values = c('grey50','orange','red'),
                     limits = c('control','wet','dry'),
                     labels = c('Unburned\ncontrol','Moist soil\nburn','Dry soil\nburn')) +
  scale_x_discrete(limits = c('O','A'),
                     labels = c('Organic','Mineral')) +
  #scale_fill_manual(values = c("deepskyblue3","saddlebrown"),
  # scale_fill_manual(values = c("slateblue4","palegreen3"),
  # #scale_fill_manual(values = c("darkgreen","saddlebrown"),
  #                    limits = c('O','A'),
  #                    labels = c('Organic','Mineral')) +
  # scale_x_discrete(limits = c('control','wet','dry'),
  #                    labels = c('Unburned\ncontrol','Wet soil\nburn','Dry soil\nburn')) +
  theme(axis.text.y = element_text(size = 14),
        legend.title = element_text(size=12),
        axis.text.x = element_text(size = 14),
        axis.title.x = element_text(size = 14, margin = margin(t=10, b=10)),
        axis.title.y = element_text(size = 14, margin = margin(r=10, l=10)),
        strip.text = element_text(size=14),
        legend.text = element_text(size=12),
        legend.position = '',
       # legend.background = element_rect(fill = "darkgray"),
        legend.box = "vertical")

cowplot::plot_grid(pChange_pH, pChange_CN, ncol=2)

# Stats ----
df.stats <- subset(df.meta, incub.trtmt == 'SI' )

test = aov(C.N.ratio ~ burn.trtmt*horizon, subset(df.stats))
summary(test)
TukeyHSD(test)

# Respiration coefficients ----
#   Decay model coefficients at end of 5 week incubation (Days=34)
#   after adjusting for total C
df.coefs <- read.csv('../../data/2-pool-decay-model-output-SI-adjusted-for-total-C.csv') %>%
  subset(t==34) 
# df.coefs <- read.csv('../../data/2-pool-decay-model-output-LIwA-adjusted-for-total-C.csv') %>%
#   subset(t==180)

df.stack <- gather(df.coefs, 
                   key = 'coefs', 
                   value = value,M1,M2,k1,k2)

df.stack$coefs <- factor(df.stack$coefs, levels = c('M1', 'M2', 'k1', 'k2'))

pResp2 = ggplot(subset(df.stack, coefs %in% c('M1','M2')), aes(x=burn.trtmt)) + 
  geom_boxplot(aes(y=value, fill=burn.trtmt),varwidth = FALSE)+
  #geom_boxplot(aes(y=log(M2), fill=burn.trtmt))+
  theme_bw()+
  labs(x= 'Burn treatment',
     #  y = 'Decay rate constant for fast C pool \nlog(k1)',
       y = 'Fractional pool size',
       fill = 'Burn treatment',
       color = 'Burn treatment') + 
  scale_fill_manual(values = c('grey50','orange','red'),
                     limits = c('control','wet burn','dry burn'),
                     labels = c('Unburned\ncontrol','Moist soil\nburn','Dry soil\nburn')) +
  scale_x_discrete(limits = c('control','wet burn','dry burn'),
                     labels = c('Unburned\ncontrol','Moist soil\nburn','Dry soil\nburn')) +
  #scale_x_discrete(limits = c('O','A'), 
  #                 labels = c('Organic','Mineral'))
  facet_wrap(~coefs, ncol=1,scales = 'free_y',
             labeller = labeller(coefs = CoefsLabel))+
  theme(axis.text.y = element_text(size = 12),
        legend.title = element_text(size=12),
        axis.text.x = element_text(size = 12),
        axis.title.x = element_text(size = 12, margin = margin(t=10, b=10)),
        axis.title.y = element_text(size = 14, margin = margin(r=25, l=10)),
        strip.text = element_text(size=12),
        legend.text = element_text(size=12),
       # legend.background = element_rect(fill = "darkgray"),
        legend.box = "vertical", 
       legend.position = '',
       strip.background = element_rect(colour="black", fill="grey92"))

pResp1 = ggplot(subset(df.stack, coefs %in% c('k1','k2')), aes(x=burn.trtmt)) + 
  geom_boxplot(aes(y=value, fill=burn.trtmt),varwidth = FALSE)+
  #geom_boxplot(aes(y=log(M2), fill=burn.trtmt))+
  theme_bw()+
  labs(x= 'Burn treatment',
     #  y = 'Decay rate constant for fast C pool \nlog(k1)',
       y = 'Decay rate',
       fill = 'Burn treatment',
       color = 'Burn treatment') + 
  scale_fill_manual(values = c('grey50','orange','red'),
                     limits = c('control','wet burn','dry burn'),
                     labels = c('Unburned\ncontrol','Moist soil\nburn','Dry soil\nburn')) +
  scale_x_discrete(limits = c('control','wet burn','dry burn'),
                     labels = c('Unburned\ncontrol','Moist soil\nburn','Dry soil\nburn')) +
  #scale_x_discrete(limits = c('O','A'), 
  #                 labels = c('Organic','Mineral'))
  facet_wrap(~coefs, ncol=1,scales = 'free_y',
             labeller = labeller(coefs = CoefsLabel))+
  theme(axis.text.y = element_text(size = 12),
        legend.title = element_text(size=12),
        axis.text.x = element_text(size = 12),
        axis.title.x = element_text(size = 12, margin = margin(t=10, b=10)),
        axis.title.y = element_text(size = 14, margin = margin(r=10, l=10)),
        strip.text = element_text(size=12),
        legend.text = element_text(size=14),
       # legend.background = element_rect(fill = "darkgray"),
        legend.box = "vertical", 
       legend.position = '',
       strip.background = element_rect(colour="black", fill="grey92"))

cowplot::plot_grid(pResp2, pResp1, ncol=2)

# Stats ----
df.stats <- subset(df.meta, incub.trtmt != 'pb' & horizon == 'O')

summary(subset(df.stats, incub.trtmt == 'LIwA' & texture != 'organic' &
                 horizon == 'O' & burn.trtmt == 'wet')$k1)
sd(subset(df.stats, incub.trtmt == 'LIwA' & burn.trtmt == 'wet')$k1)

test = aov(k2 ~ burn.trtmt, subset(df.stats, incub.trtmt == 'LIwA' & texture != 'organic'))
summary(test)
TukeyHSD(test)

test = wilcox.test(M1~burn.trtmt, data = subset(df.stats, incub.trtmt == 'SI' & burn.trtmt != 'control'))
test

test = aov(M1 ~ texture, data = subset(df.meta, incub.trtmt %in% c('LIwA','LIn') & horizon == 'O'))

test

p = ggplot(df.stats, aes(x=burn.trtmt, y =k1)) + 
  geom_boxplot() + 
  facet_grid(~incub.trtmt) +
  theme_bw()

```

# Figure 3: RNA and DNA concentration (ng/g soil)
```{r}
df <- read.csv('../../data/raw-data/pb-RNA-and-DNA-conc.csv')

colnames(df)[1] <- 'core.id.hor.incub'

df.sub <- df.meta %>%
  subset(incub.trtmt == 'pb') %>%
  subset(select = c(core.id.hor.incub, DNA.type, horizon, burn.trtmt))

df.DNA <- subset(df, DNA.type == 'gDNA') %>%
  mutate(Conc.DNA = Concentration_ng.uL) %>%
  subset(select = c('core.id.hor.incub', 'Conc.DNA'))
df.RNA <- subset(df, DNA.type == 'cDNA') %>%
  mutate(Conc.RNA = Concentration_ng.uL) %>%
  subset(select = c('core.id.hor.incub', 'Conc.RNA'))

df <- merge(df, df.sub)

NaLabels = c('RNA','DNA')
names(NaLabels) <- c('cDNA','gDNA')

p1 = ggplot(df) + 
  geom_boxplot(aes(x=horizon, y = nucleic.acid.conc.in.soil.ng.per.g/1000, fill = burn.trtmt),varwidth = FALSE) + 
  theme_bw()+
  facet_wrap(~DNA.type, ncol=2, labeller = labeller(DNA.type = NaLabels))+
  labs(x= 'Soil horizon',
       #  y = 'Decay rate constant for fast C pool \nlog(k1)',
       y = 'Concentration\n(ug/g soil)',
       fill = 'Burn treatment') + 
  scale_fill_manual(values = c('grey50','orange','red'),
                     limits = c('control','wet','dry'),
                     labels = c('Unburned\ncontrol','Moist soil\nburn','Dry soil\nburn')) +
  scale_x_discrete(limits = c('O','A'),
                     labels = c('Organic','Mineral')) +
  # scale_fill_manual(values = c("darkgreen","grey70"),
  #                   limits = c('O','A'),
  #                   labels = c('Organic','Mineral')) +
  # scale_x_discrete(limits = c('control','wet','dry'),
  #                  labels = c('Unburned\ncontrol','Moist soil\nburn','Dry soil\nburn')) +
  theme(axis.text.y = element_text(size = 14),
        legend.title = element_text(size=14),
        axis.text.x = element_text(size = 13),
        axis.title.x = element_text(size = 14, margin = margin(t=10, b=10)),
        axis.title.y = element_text(size = 14, margin = margin(r=10, l=10)),
        strip.text = element_text(size=14),
        legend.text = element_text(size=14),
        legend.position = '',
        # legend.background = element_rect(fill = "darkgray"),
        legend.box = "vertical")
p1

p_DNA = ggplot(subset(df, DNA.type == 'gDNA')) + 
  geom_boxplot(aes(x=horizon, y = nucleic.acid.conc.in.soil.ng.per.g/1000, fill = burn.trtmt),varwidth = FALSE) + 
  theme_bw()+
  labs(x= 'Soil horizon',
       #  y = 'Decay rate constant for fast C pool \nlog(k1)',
       y = 'DNA concentration\n(ug/g soil)',
       fill = 'Burn treatment') + 
  scale_fill_manual(values = c('grey50','orange','red'),
                     limits = c('control','wet','dry'),
                     labels = c('Unburned\ncontrol','Moist soil\nburn','Dry soil\nburn')) +
  scale_x_discrete(limits = c('O','A'),
                     labels = c('Organic','Mineral')) +
  # scale_fill_manual(values = c("darkgreen","grey70"),
  #                   limits = c('O','A'),
  #                   labels = c('Organic','Mineral')) +
  # scale_x_discrete(limits = c('control','wet','dry'),
  #                  labels = c('Unburned\ncontrol','Moist soil\nburn','Dry soil\nburn')) +
  theme(axis.text.y = element_text(size = 14),
        legend.title = element_text(size=14),
        axis.text.x = element_text(size = 13),
        axis.title.x = element_text(size = 14, margin = margin(t=10, b=10)),
        axis.title.y = element_text(size = 14, margin = margin(r=10, l=10)),
        strip.text = element_text(size=14),
        legend.text = element_text(size=14),
        legend.position = '',
        # legend.background = element_rect(fill = "darkgray"),
        legend.box = "vertical")

p_RNA = ggplot(subset(df, DNA.type == 'cDNA')) + 
  geom_boxplot(aes(x=horizon, y = nucleic.acid.conc.in.soil.ng.per.g/1000, fill = burn.trtmt),varwidth = FALSE) + 
  theme_bw()+
  labs(x= 'Soil horizon',
       #  y = 'Decay rate constant for fast C pool \nlog(k1)',
       y = 'RNA concentration\n(ug/g soil)',
       fill = 'Burn treatment') + 
  scale_fill_manual(values = c('grey50','orange','red'),
                     limits = c('control','wet','dry'),
                     labels = c('Unburned control','Moist soil burn','Dry soil burn')) +
  scale_x_discrete(limits = c('O','A'),
                     labels = c('Organic','Mineral')) +
  # scale_fill_manual(values = c("darkgreen","grey70"),
  #                   limits = c('O','A'),
  #                   labels = c('Organic','Mineral')) +
  # scale_x_discrete(limits = c('control','wet','dry'),
  #                  labels = c('Unburned\ncontrol','Moist soil\nburn','Dry soil\nburn')) +
  theme(axis.text.y = element_text(size = 14),
        legend.title = element_text(size=14),
        axis.text.x = element_text(size = 13),
        axis.title.x = element_text(size = 14, margin = margin(t=10, b=10)),
        axis.title.y = element_text(size = 14, margin = margin(r=10, l=10)),
        strip.text = element_text(size=14),
        legend.text = element_text(size=14),
        legend.position = '',
        # legend.background = element_rect(fill = "darkgray"),
        legend.box = "vertical")

cowplot::plot_grid(p_DNA, p_RNA, ncol=2)

test <- aov(nucleic.acid.conc.in.soil.ng.per.g ~ burn.trtmt*horizon, data = subset(df, DNA.type == 'gDNA'))
summary(test)
TukeyHSD(test)

```

Bray-Curtis dissimilarity?
```{r, message = FALSE}
# This is a possibly convoluted piece of code to first create a Bray-Curtis
#     dissimilarity matrix and then pull out the dissimilarities between dry/wet
#     burns and corresponding control cores.

# First: Calculate Bray-curtis dissimilarity matrix ----
ps.cDNA.raw <- prune_samples(sample_data(ps.raw.full)$DNA.type == 'cDNA', ps.raw.full)
ps.cDNA.raw <- prune_taxa(taxa_sums(ps.cDNA.raw) > 0, ps.cDNA.raw)

dist.bray.full <- as.matrix(distance(ps.raw.full, method = 'bray', type = 'samples'))

# Look at some figures
hist(dist.bray.full)
heatmap(dist.bray.full)

# Convert distance matrix into dataframe
df <- data.frame(dist.bray.full)

# Bring in meta data and subset for specific analysis
df.meta.sub <- subset(df.meta, select = c(Full.id, burn.trtmt,site, DNA.type,
                                          incub.trtmt,core, horizon,
                                          pH, change.pH))

# Create a dataframe with the colnames and Full.ID
x <- data.frame(str_split(colnames(df)[1:460], 'X', simplify = TRUE)[,2])

# Rename columes:
colnames(x)[1] <- 'ID.bray'

# Clean up x. 
x <- x %>%
  mutate(Full.id = ID.bray)

x$Full.id <- str_replace_all(x$Full.id, "\\.", "-")

x$X <- 'X'

x <- x %>%
  unite(Bray.ID, c('X',"ID.bray"), sep = '', remove = TRUE)

# Next, pull out the control cores from the meta data and assign them to "Control.id"
df.meta.control <- df.meta %>%
  subset(burn.trtmt == 'control') %>%
  mutate(Control.id = Full.id) %>%
  subset(select = c(Control.id, site, horizon, incub.trtmt, DNA.type))

# Combine list of control cores with full df.meta
df.meta.list <- merge(subset(df.meta.sub, select = c(Full.id, site, horizon, 
                                                     DNA.type,burn.trtmt, incub.trtmt, 
                                                     pH, change.pH)), 
                      df.meta.control, by = c('site', 'horizon', 'incub.trtmt', 'DNA.type'))

# ANd then combine the Full.id list with the control core meta data I just created.
#     df.list now contains Full.id with the corresponding control sample for each 
#     dry/wet burn and a colume ("Bray.ID") that has the sample ID formatted to 
#     match the distance matrix names.
df.list <- x %>%
  merge(df.meta.list, by = 'Full.id') %>%
  subset(burn.trtmt != 'control')

df.list$burn.trtmt <- as.character(df.list$burn.trtmt)
df.list$horizon <- as.character(df.list$horizon)
df.list$incub.trtmt <- as.character(df.list$incub.trtmt)

# Create and empty data frame to store distances
df.distances <- data.frame("Full.id" = df.list$Full.id, 
                           'control.core' = 'core',
                           #'burn.trtmt' = 'burn', 'horizon' = 'H',
                           #'DNA.type' = 'DNA.type',
                           'BC.distance.btw.burned.and.control' = 'distance'
                           #'incub.trtmt' = 'incub', 'pH' = 0
                           ) %>%
  unique()

# Now run a convoluted loop:
for (i in 1:nrow(df.list)) { # For each sample (row in df.list)
  for (j in 1:nrow(df)) { # For each row in the dissimilarity matrix
    for (k in 1:ncol(df)){ # And for each colume in the dissimilarity matrix
      if(colnames(df)[k] == df.list$Bray.ID[i] &  # If the distance matrix colume ID matches the sample ID in the list
         rownames(df)[j] == df.list$Control.id[i]) { # And the rowname ID matches the corresponding control core, then...
        df.distances$BC.distance.btw.burned.and.control[i] = df[j,k] # Pull out the distance and store it in new dataframe
        #df.distances$burn.trtmt[i] = df.list$burn.trtmt[i]
        df.distances$control.core[i] = df.list$Control.id[i]
        #df.distances$horizon[i] = df.list$horizon[i]
        #df.distances$incub.trtmt[i] = df.list$incub.trtmt[i]
        #df.distances$DNA.type[i] = df.list$DNA.type[i]
        #df.distances$pH[i] = df.list$pH[i]
        #df.distances$change.pH[i] = df.list$change.pH[i]
      }
    } 
  } 
}

df.distances$BC.distance.btw.burned.and.control <- as.numeric(df.distances$BC.distance.btw.burned.and.control)

# Save backup dataframe.
df.final <- df.distances

# -----
df.distances <- read.csv('../data/bray-curtis-dissimilarites.csv')

df.distances <- merge(df.distances, subset(df.meta, burn.trtmt != 'control'))                            
dim(df.distances)

# Plots ----
p=ggplot(subset(df.distances, DNA.type == 'cDNA')) + 
  geom_boxplot(aes(x=burn.trtmt, y = BC.distance.btw.burned.and.control, fill = horizon),
               alpha = 0.7) +
  theme_bw() +
  facet_wrap(~incub.trtmt, ncol=4,
             labeller = labeller(incub.trtmt = IncubLabels)) + 
  labs(x='Burn treatment',
       y = 'Bray-Curtis dissimilarity \ncompared to control', 
       fill = 'Soil horizon') + 
  scale_fill_manual(values = c('black','grey40'),
                     limits = c('O','A'),
                     labels = c('Organic','Mineral')) +
  scale_x_discrete(limits = c('wet','dry'),
                   labels = c('Moist soil \nburn','Dry soil \nburn')) +
  theme(axis.title = element_text(size=18),
        axis.text = element_text(size=18),
        legend.title = element_text(size=18),
        legend.text = element_text(size=18),
        strip.text = element_text(size=18)) + 
  ylim(0,1)

test <- aov(BC.distance.btw.burned.and.control~burn.trtmt*horizon, data = subset(df.distances, DNA.type == 'cDNA'))
summary(test)
TukeyHSD(test)

```

Richness estimates?

If the community composition data of a single sample includes few singletons, we assume that sequencing captured most of the taxa present in the sample. If the data includes many singletons, we assume that sequencing missed many rare taxa. We can use breakaway to estiamte how many taxa were missed and use this to estimate richness. 
```{r}

# Load data and calculate richness ----
df.pb.gDNA.OTU <- read.csv('../data/sequence-data/LibCombined/phyloseq-objects/df.pb.gDNA.norm.csv')
df.pb.cDNA.OTU <- read.csv('../data/sequence-data/LibCombined/phyloseq-objects/df.pb.cDNA.norm.csv')
df.SI.OTU <- read.csv('../data/sequence-data/LibCombined/phyloseq-objects/df.SI.norm.csv')
df.LI.norm <- read.csv('../data/sequence-data/LibCombined/phyloseq-objects/df.LI.norm.csv')

head(df.pb.cDNA.OTU)

df.counts.pb.cDNA <- df.pb.cDNA.OTU %>%
  group_by(Full.id) %>%
  mutate(counts = length(unique(OTU))) %>%
  subset(select = c(Full.id, burn.trtmt, horizon, DNA.type, incub.trtmt, counts)) %>%
  unique()

df.counts.pb.gDNA <- df.pb.gDNA.OTU %>%
  group_by(Full.id) %>%
  mutate(counts = length(unique(OTU))) %>%
  subset(select = c(Full.id, burn.trtmt, horizon, DNA.type, incub.trtmt, counts)) %>%
  unique()

df.counts.SI <- df.SI.OTU %>%
  group_by(Full.id) %>%
  mutate(counts = length(unique(OTU))) %>%
  subset(select = c(Full.id, burn.trtmt, horizon, DNA.type, incub.trtmt, counts)) %>%
  unique()

df.counts.LI <- df.LI.norm %>%
  group_by(Full.id) %>%
  mutate(counts = length(unique(OTU))) %>%
  subset(select = c(Full.id, burn.trtmt, horizon, DNA.type, incub.trtmt, counts)) %>%
  unique()

df.counts <- rbind(df.counts.pb.cDNA, df.counts.pb.gDNA,
                   df.counts.SI, df.counts.LI)

df.counts$burn.trtmt <-  factor(df.counts$burn.trtmt, levels = c('control','wet','dry'))
df.counts$horizon <-  factor(df.counts$horizon, levels = c('O', 'A'))

df.counts.pb <- subset(df.counts, incub.trtmt == 'pb')

# Plots ----
p = ggplot(df.counts.pb, aes(x=burn.trtmt, y=counts)) +
  geom_boxplot(aes(fill = burn.trtmt), alpha = 0.7) + 
  geom_point(color = 'black', alpha = 0.4) + 
  theme_bw() + 
  facet_grid(horizon~incub.trtmt+DNA.type, labeller = labeller(incub.trtmt = IncubLabels,
                                                               horizon = HorizonLabels, 
                                                               DNA.type = DNALabels)) + 
  labs(y = 'unique non-zero OTU count')

test <- aov(counts ~ burn.trtmt, data = subset(df.counts.pb.gDNA, horizon =='O'))
summary(test)
TukeyHSD(test)

# Plot richness using "plot_richness" function in phyloseq: ----
ps <- prune_samples(sample_data(ps.raw.full)$incub.trtmt == 'pb',
                    ps.raw.full)

sample_data(ps)$burn.trtmt <- factor(sample_data(ps)$burn.trtmt, levels = c('control','wet','dry'))
sample_data(ps)$horizon <- factor(sample_data(ps)$horizon, levels = c('O','A'))

p = plot_richness(ps, x = "burn.trtmt", measures = 'Shannon') +
  theme_bw() + 
  geom_boxplot(aes(fill = horizon), alpha = 0.6) + 
  facet_grid(horizon~DNA.type+incub.trtmt, labeller = labeller(incub.trtmt = IncubLabels,
                                                   DNA.type = DNALabels,
                                                      horizon = HorizonLabels))
```

RNA:DNA ratios?
```{r}
df.norm.ratios <- read.csv('../data/sequence-data/LibCombined/phyloseq-objects/RNA_DNA_norm_ratios.csv')
# RNA:DNA ratios were calculated by (RNA rel. abundance)/(DNA rel. abundance)
#     In instances of DNA rel. abundance = 0, I set DNA abundance = 0.0000038, which
#     is equal to the lowest measured rel. abundance of DNA across the dataset.
#     In instances of RNA re. abundance = 0, I set ratio = 0.

df.norm.ratios$horizon <- factor(df.norm.ratios$horizon, levels = c('O','A'))
df.norm.ratios$burn.trtmt <- factor(df.norm.ratios$burn.trtmt, levels = c('control','wet','dry'))

# Plot 
p = ggplot(data = subset(df.norm.ratios, DNA.abundance > 0.001), aes(x=burn.trtmt, y = RNA.DNA.ratio, fill = horizon)) + 
  geom_boxplot()+ 
  #stat_smooth(method=lm,se = FALSE,color='black') +
  #facet_wrap(~horizon, ncol=1,scale = 'free')+
  theme_bw() + 
  labs(title = 'RNA:DNA \n*only taxa present in paired burned/unburned trtmts')

# What fraction of total reads for each sample fall below 0.001 relative abundance threshold? -----
df.DNA.tally <- df.norm.ratios %>%
  subset(RNA.abundance + DNA.abundance > 0) %>%
  #subset(DNA.abundance > 0) %>%
  group_by(core.id.hor.incub, burn.trtmt, horizon) %>%
  mutate(counts = length(OTU)) %>%
  group_by(core.id.hor.incub, burn.trtmt, horizon,counts) %>%
  tally(DNA.abundance > 0.001) %>%
  #mutate(DNA.fraction.above.cutoff = n/counts)
  mutate(non.zero.DNA.fraction = n/counts)
head(df.DNA.tally)

df.RNA.tally <- df.norm.ratios %>%
  subset(RNA.abundance + DNA.abundance > 0) %>%
  #subset(RNA.abundance > 0) %>%
  group_by(core.id.hor.incub, burn.trtmt, horizon) %>%
  mutate(counts = length(OTU)) %>%
  group_by(core.id.hor.incub, burn.trtmt, horizon,counts) %>%
  tally(RNA.abundance > 0.001)%>%
  mutate(non.zero.RNA.fraction = n/counts)
  #mutate(RNA.fraction.above.cutoff = n/counts)
head(df.RNA.tally)

df.DNA.tally$burn.trtmt <- factor(df.DNA.tally$burn.trtmt, levels = c('control','wet','dry'))
df.RNA.tally$burn.trtmt <- factor(df.RNA.tally$burn.trtmt, levels = c('control','wet','dry'))

p = ggplot(df.DNA.tally, aes(x=burn.trtmt, y =DNA.fraction.above.cutoff, fill = horizon)) + 
  geom_boxplot() + theme_bw()

test <- aov(DNA.fraction.above.cutoff ~ burn.trtmt*horizon, data = df.DNA.tally)
summary(test)
TukeyHSD(test)

test <- aov(non.zero.RNA.fraction ~ burn.trtmt*horizon, data = df.RNA.tally)
summary(test)
TukeyHSD(test)

# Re-calculate ratio including only taxa present in both burned and unburned cores: ----
df.burned <- subset(df.norm.ratios, burn.trtmt == 'dry') %>%
  subset(DNA.abundance > 0) %>%
  mutate(burned.DNA.abundance = DNA.abundance) %>%
  subset(select = c(site, horizon, OTU, burned.DNA.abundance))

df.control <- subset(df.norm.ratios, burn.trtmt == 'control') %>%
  subset(DNA.abundance > 0) %>%
  mutate(unburned.DNA.abundance = DNA.abundance) %>%
  subset(select = c(site, horizon, OTU, unburned.DNA.abundance))

df.shared.OTUs <- merge(df.burned, df.control)

df.common.taxa <- subset(df.norm.ratios, burn.trtmt != 'wet') %>%
  merge(df.shared.OTUs) %>%
  mutate(ratio = RNA.abundance/DNA.abundance) %>%
  group_by(core.id.hor.incub) %>%
  mutate(mean.ratio = mean(ratio))

df.counts <- df.common.taxa %>%
  group_by(core.id.hor.incub, burn.trtmt, horizon) %>%
  tally(DNA.abundance > 0)

dim(df.common.taxa)

# Plot 
p = ggplot(data = df.common.taxa, aes(x=burn.trtmt, y = ratio, fill = horizon)) + 
  geom_boxplot()+ 
  #stat_smooth(method=lm,se = FALSE,color='black') +
  #facet_wrap(~horizon, ncol=1,scale = 'free')+
  theme_bw() + 
  labs(title = 'RNA:DNA \n*only taxa present in paired burned/unburned trtmts')

test <- aov(ratio ~ burn.trtmt*horizon, data = df.common.taxa)
summary(test)
TukeyHSD(test)

# Recalculating ratios using raw reads: -----
ps.raw.gDNA <- prune_samples(sample_data(ps.raw.pb)$DNA.type == 'gDNA', ps.raw.pb)
ps.raw.gDNA
ps.raw.gDNA <- prune_taxa(taxa_sums(ps.raw.gDNA)>0, ps.raw.gDNA)
ps.raw.gDNA

ps.raw.cDNA <- prune_taxa(taxa_sums(ps.raw.cDNA)>0, ps.raw.cDNA)

df.raw.pb.cDNA <- psmelt(ps.raw.cDNA)
df.raw.pb.gDNA <- psmelt(ps.raw.gDNA)

df.RNA <- df.raw.pb.cDNA %>%
  mutate(RNA.abundance = Abundance) %>%
  subset(select = c(core.id.hor.incub, OTU, RNA.abundance))

df.DNA <- df.raw.pb.gDNA %>%
  mutate(DNA.abundance = Abundance) %>%
  subset(select = c(core.id.hor.incub, OTU, DNA.abundance, horizon, site, pH, hrzn.degree.hrs,burn.trtmt))

rownames(df.RNA) <- NULL
rownames(df.DNA) <- NULL

df.merge <- merge(df.RNA, df.DNA, by = c("core.id.hor.incub", "OTU"), all = TRUE)

df.pb <- df.merge %>%
  group_by(site, OTU) %>%
  mutate(min = min(DNA.abundance)) %>%
  subset(min > 0)

df.pb[is.na(df.pb$RNA.abundance),3] <- 0

df.pb <- df.pb %>%
  mutate(ratio = RNA.abundance/DNA.abundance)

# Plot
df.pb$burn.trtmt <- factor(df.pb$burn.trtmt, levels = c('control','wet','dry'))

p = ggplot(data = subset(df.pb,horizon =='O'), aes(x=burn.trtmt, y = (ratio), color = burn.trtmt)) + 
  geom_boxplot()+ 
  #stat_smooth(method=lm,se = FALSE,color='black') +
  facet_wrap(~horizon, ncol=1,scale = 'free')+
  theme_bw() + 
  labs(title = 'RNA:DNA \nonly taxa present across all trtmts')

test <- aov((ratio) ~ burn.trtmt, data = subset(df.pb, horizon == 'O'))
summary(test)
TukeyHSD(test)

# Alternative calculations for RNA:DNA ratios: ----

### Set ratio = 100 if DNA rel. abundance = 0 (Bowsher et al. 2019)
for (i in 1:nrow(df.norm.ratios)) {
  if(df.norm.ratios$DNA.abundance[i] == 0) {
    df.norm.ratios$RNA.DNA.ratio[i] = 100
  } else {
    df.norm.ratios$RNA.DNA.ratio[i] = df.norm.ratios$RNA.DNA.ratio[i]
  }
}

### If DNA abundance = 0, change it to 1.  (Bowsher et al. 2019)
#     But this would only work for raw count data.
for (i in 1:nrow(df.norm.ratios)) {
  if(df.norm.ratios$DNA.abundance[i] == 0) {
    df.norm.ratios$DNA.abundance[i] = 0.000038
  } else {
    df.norm.ratios$DNA.abundance[i] = df.norm.ratios$DNA.abundance[i]
  }
}
df.norm.ratios$RNA.DNA.ratio = df.norm.ratios$RNA.abundance/df.norm.ratios$DNA.abundance

### Add 1 to all DNA and RNA abundances (Bowsher et al. 2019)
df.norm.ratios$DNA.abundance = df.norm.ratios$DNA.abundance + 0.000038

#df.norm.ratios$RNA.abundance = df.norm.ratios$RNA.abundance + 0.000038

df.norm.ratios <- subset(df.norm.ratios, DNA.abundance > 0.0001 || RNA.abundance >0.0001 )

df.norm.ratios$RNA.DNA.ratio = df.norm.ratios$RNA.abundance/df.norm.ratios$DNA.abundance
```

# Figure 4 - weighted mean gene copy number
Which bacteria are fast-growing following burn treatments of varying severity and under a range of moisture conditions?
```{r}
df.WMCN.pb <- read.csv('../../data/WtMeanCopyNum_pb_gDNA.csv')
df.WMCN.SI <- read.csv('../../data/WtMeanCopyNum_SI.csv')
df.WMCN.LIwA <- read.csv('../../data/WtMeanCopyNum_LIwA.csv')

df.WMCN.SI <- subset(df.WMCN.SI, core.id != '19UW-WB-11-02' & 
                    Full.id != '19UW-WB-06-08-A-SI-duplicate' &
                    Full.id != '19UW-WB-07-02-O-SI-duplicate' &
                    Full.id != '19UW-WB-11-03-O-SI-duplicate'&
                    Full.id != '19UW-WB-19-07-A-SI-duplicate' &
                    Full.id != '19UW-WB-08-10-O-SI' )

summary(subset(df.WMCN.SI, burn.trtmt == 'control' &horizon == 'O')$WtMeanCopyNum)
sd(subset(df.WMCN.SI, burn.trtmt == 'control' &horizon == 'O')$WtMeanCopyNum)
summary(subset(df.WMCN.SI, burn.trtmt == 'control' &horizon == 'A')$WtMeanCopyNum)
sd(subset(df.WMCN.SI, burn.trtmt == 'control' &horizon == 'A')$WtMeanCopyNum)

test = aov(WtMeanCopyNum~ burn.trtmt*horizon, data = df.WMCN.SI)
summary(test)
TukeyHSD(test)

df.WMCN <- rbind(df.WMCN.SI, df.WMCN.LIwA, df.WMCN.pb)

df.WMCN <- df.WMCN %>%
  unite(core.id.hor, c(core.id, horizon), sep = '-', remove=FALSE)

df.WMCN <- merge(df.WMCN, df.meta)

# CHeck for normality:
p.QQ = qqplot(df.WMCN$hrzn.degree.hrs, df.WMCN$WtMeanCopyNum)

# Plot ----
df.WMCN$burn.trtmt <- factor(df.WMCN$burn.trtmt, levels = c('control','wet','dry'))
df.WMCN$horizon <- factor(df.WMCN$horizon, levels = c('O','A'))
df.WMCN$incub.trtmt <- factor(df.WMCN$incub.trtmt, levels = c('pb','SI','LIwA'))

p1 = ggplot(data = subset(df.WMCN, incub.trtmt != 'LIwA')) + 
  geom_boxplot(aes(x=burn.trtmt, y=WtMeanCopyNum, fill=horizon),alpha=0.7)+
  facet_wrap(~incub.trtmt, labeller = labeller(incub.trtmt = IncubLabels))+
  theme_bw() + 
  scale_fill_manual(values = c("darkgreen","grey70"),
                     labels = c('Organic','Mineral'),
                     limits = c('O','A'))+
  scale_x_discrete(limits = c('control','wet','dry'),
                     labels = c('Unburned\ncontrol','Moist soil\nburn','Dry soil\nburn')) +
  labs(x='Burn treatment',
       y='Weighted mean predicted \n16S rRNA gene copy number',
       fill = 'Soil horizon') +
  theme(axis.text.y = element_text(size = 12),
        axis.text.x = element_text(size = 12),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14),
        strip.text = element_text(size=14),
        legend.position = 'right',
        axis.title.y = element_text(size = 14, margin = margin(r=10, l=10)),
        axis.title.x = element_text(size = 14, margin = margin(t=10, b=10))) +
  ylim(1,NA)

# Stats
df.WMCN.O <- subset(df.WMCN, horizon == 'O')
df.WMCN.A <- subset(df.WMCN, horizon == 'A')

fit.df.WMCN.O <- lm(WtMeanCopyNum~hrzn.degree.hrs, data=df.WMCN.O)
fit.df.WMCN.A <- lm(WtMeanCopyNum~hrzn.degree.hrs, data=df.WMCN.A)

summary(fit.df.WMCN.O)
# p = 2*10^-16, R^2 adj = 0.6081
summary(fit.df.WMCN.A)
# p = 2*10^-16, R^2 adj = 0.7883

# Weighted mean copy number STATS:
test <- aov(WtMeanCopyNum ~ burn.trtmt*horizon, data = subset(df.WMCN, incub.trtmt == 'SI'))
summary(test)
TukeyHSD(test)

test.sum <- df.WMCN %>%
  group_by(burn.trtmt, horizon) %>%
  summarize(mean.WMCN = mean(WtMeanCopyNum))

# Calculate WMCN change from pb to SI: ----
df.WMCN.initial <- subset(df.WMCN, incub.trtmt == 'pb') %>%
  mutate(WMCN.pb = WtMeanCopyNum) %>%
  subset(select = c(core.id, horizon, burn.trtmt, WMCN.pb))

df.WMCN.SI <- df.WMCN.SI %>%
  merge(df.WMCN.initial, all = TRUE) %>%
  mutate(change.WMCN = WtMeanCopyNum - WMCN.pb)

df.WMCN.SI$burn.trtmt <- factor(df.WMCN.SI$burn.trtmt, levels = c('control','wet','dry'))

p = ggplot(df.WMCN.SI, aes(x=burn.trtmt, y = change.WMCN, fill = horizon)) + 
  geom_boxplot() + 
  geom_hline(yintercept = 0)+
  theme_bw() + 
  labs(x = 'Comparison between samples 24 h and 5 wk post-burn',
       y = 'Change in predicted wt. mean gene copy number')

test <- aov(change.WMCN ~ burn.trtmt*horizon, data = df.WMCN.SI)
summary(test)
TukeyHSD(test)

# Compare copy number of positive/negative responders: ----
df.full.OTU <- read.csv('../../data/sequence-data/LibCombined/corncob-output/Manuscript/Trait-responder-OTUs.csv')
RDP <- read.csv('../../data/All-OTUs-CopyNum.csv')

df.fast <- subset(df.full.OTU, experiment == 'fast growth' &
                        Estimate > (mean(Estimate)-1*sd(Estimate))) %>%
  mutate(fast.est = Estimate)%>%
  dplyr::group_by(OTU) %>%
  dplyr::mutate(max.Estimate = max(Estimate)) %>%
  subset(Estimate == max.Estimate) %>%
  subset(select = c(OTU, Estimate, SE, t, p, p_fdr,Comparison, experiment,
                    horizon, Domain, Phylum, Class, Order, Family, Genus, 
                    Species, DNA.type, fast.est)) %>%
  unique()

ps.norm.SI <- prune_samples(sample_data(ps.norm.full)$incub.trtmt == 'SI', ps.norm.full)
ps.norm.SI <- prune_taxa(taxa_sums(ps.norm.SI)>0, ps.norm.SI)

df.not.responders <- data.frame(taxa_names(ps.norm.SI))

df.not.responders <- subset(df.not.responders, !(taxa_names.ps.norm.SI. %in% df.responders$OTU))

df.not.responders$Responders = 'no'
df.responders <- subset(df.fast, select = c(OTU)) %>%
  mutate(Responders = 'yes')
colnames(df.not.responders)[1] = 'OTU'

df.responders <- rbind(df.responders, df.not.responders) %>%
  merge(RDP)

p = ggplot(df.responders, aes(x=Responders, y = CopyNum)) + 
  geom_boxplot() + 
  theme_bw()

df.SI.responders <- df.full.OTU %>%
  subset(experiment == 'fast growth') %>%
  merge(RDP)

dim(df.SI.responders)

df.SI.positive = subset(df.SI.responders, Estimate > 0 ) %>%
  group_by(OTU) %>%
  mutate(Estimate = max(Estimate)) %>%
  mutate(Response = 'positive') %>%
  subset(select = c(OTU, Estimate, Comparison, horizon, Domain, Phylum, Class, Order, 
                    Family, Genus, Species, GenusRDP, CopyNum, Response)) %>%
  unique()

df.SI.negative = subset(df.SI.responders, Estimate < 0 ) %>%
  group_by(OTU) %>%
  mutate(Estimate = min(Estimate)) %>%
  mutate(Response = 'negative') %>%
  subset(select = c(OTU, Estimate, Comparison, horizon, Domain, Phylum, Class, Order, 
                    Family, Genus, Species, GenusRDP, CopyNum, Response)) %>%
  unique()

df.SI.responders <- rbind(df.SI.positive, df.SI.negative)

p = ggplot(df.SI.responders, aes(x=Estimate, y = Class, color = Phylum)) + 
  geom_point() + 
  theme_bw() + 
  facet_wrap(~Comparison) +
  geom_vline(xintercept = 0)

df.SI.dry.responders <- subset(df.SI.responders, Comparison == 'pb.dry.v.SI.dry') %>%
  subset(!(OTU %in% subset(df.SI.responders, Comparison %in% c('pb.wet.v.SI.wet', 'pb.control.v.SI.control')))) %>%
  subset(select = c(OTU, CopyNum, Response, horizon,Estimate)) %>%
  unique()

p = ggplot(subset(df.SI.dry.responders, is.na(CopyNum) == FALSE), aes(x=Response, y = CopyNum, fill = horizon)) + 
  geom_boxplot() + 
  theme_bw() + 
  facet_wrap(~horizon)

test = aov(CopyNum ~ Response+horizon, data = df.SI.dry.responders)
summary(test)
TukeyHSD(test)

p = ggplot(subset(df.SI.dry.responders, is.na(CopyNum) == FALSE), aes(x=Estimate, y = CopyNum)) + 
  geom_point()+
  facet_grid(Response~horizon) + 
  stat_smooth(method=lm,se = FALSE,color='black') +
  theme_bw() 
```

# Figure 5 - Calculate bray-curtis dissimilarity for autoclaving
```{r, message = FALSE}
df.distances <- read.csv('../../data/bray-curtis-dissimilarites.csv')

df.distances <- merge(df.distances, subset(df.meta, burn.trtmt != 'control'))

df.distances$incub.trtmt <- factor(df.distances$incub.trtmt, levels = c('pb','SI','LIwA','LIn'))
df.distances$horizon <- factor(df.distances$horizon, levels = c('O','A'))

# plot: Dry soil burns
p3=ggplot(subset(df.distances, incub.trtmt == 'LIwA' & burn.trtmt == 'dry'),
         aes(x=change.pH, y = BC.distance.btw.burned.and.control, color = horizon, shape = horizon)) + 
  geom_point(size=3, alpha = 0.8) +
  theme_bw() +
  facet_wrap(~horizon,ncol=2,
             labeller = labeller(horizon = HorizonLabels),
             scales = 'fixed') + 
  labs(x = expression(Delta~'pH with burning'),
       y = 'Bray-Curtis dissimilarity \ncompared to control', 
       color = 'Soil horizon',
       shape = 'Soil horizon') + 
  scale_color_manual(values = c("darkgreen","grey70"),
                     limits = c('O','A'),
                     labels = c('Organic','Mineral')) +
  scale_shape_manual(values = c('circle','triangle'),
                     limits = c('O','A'),
                     labels = c('Organic','Mineral')) +
  theme(axis.title.x = element_text(size=14, margin = margin(t=10)),
        axis.title.y = element_text(size=14, margin = margin(r=10)),
        axis.text = element_text(size=12),
        legend.title = element_text(size=14),
        legend.text = element_text(size=14),
        strip.text = element_text(size=14),
        panel.spacing.x = unit(2, "lines"),
       legend.position = 'right') +
  stat_smooth(data = subset(df.distances, incub.trtmt == 'LIwA' & burn.trtmt == 'dry' & horizon == 'O'),
              method=lm,se = FALSE,color='black') +
    ylim(0,1)

test = lm(BC.distance.btw.burned.and.control~change.pH, data=subset(df.distances, incub.trtmt == 'LIwA' & horizon == 'A' & burn.trtmt == 'dry'))

test
summary(test)

# Hrzn.degree.hrs vs. Bray-Curtis dissimilarities for -pb RNA and DNA -----
p = ggplot(subset(df.meta.combined, burn.trtmt =='dry' & 
                    incub.trtmt == 'pb'),
           aes(x=hrzn.degree.hrs, y = BC.distance.btw.burned.and.control,fill = horizon)) +
  geom_point(shape='circle filled',
             stroke=1,
             alpha = 0.8,
             size=3) +
  theme_bw() +
  facet_grid(DNA.type~horizon, scales = 'free',
             labeller = labeller(incub.trtmt = IncubLabels,
                                 burn.trtmt = BurnLabels, 
                                 DNA.type = DNALabels)) + 
  labs(y = 'Bray-Curtis dissimilarity compared to control', 
       title = 'Bray-Curtis dissimilarities') + 
  scale_fill_manual(values = c('black','grey70'),
                     limits = c('O','A'),
                     labels = c('Organic','Mineral')) +
  theme(axis.title = element_text(size=12),
        axis.text = element_text(size=12),
        legend.title = element_text(size=12),
        legend.text = element_text(size=12)) + 
    stat_smooth(method=lm,se = FALSE,color='black')

# Stats -----
test <- aov(BC.distance.btw.burned.and.control ~ hrzn.degree.hrs, 
            data = subset(df.meta.combined,incub.trtmt == 'pb' & DNA.type == 'cDNA' &burn.trtmt == 'wet'))
summary(test) 
```

# Responders as fraction of total reads
```{r}
df <- read.csv('../../data/sequence-data/LibCombined/corncob-output/Manuscript/Trait-responder-OTUs.csv')

ps.responders <- prune_taxa(taxa_names(ps.norm.full) %in% df$OTU, ps.norm.full)

df.melt <- psmelt(ps.responders)

responders <- df.melt %>% group_by(Sample) %>% summarize(Abundance) 
max(responders$Abundance)
mean(responders$Abundance)

dry <- subset(df.melt, burn.trtmt == 'dry') %>% subset(select = c(Sample, horizon, burn.trtmt, Abundance, incub.trtmt, Thermo.mid.max)) %>% group_by(Sample) %>% mutate(total.Abun = sum(Abundance))

summary(dry)

# Total gDNA reads in lab experiments: 39115
# Total reads in lab experiments: 43450
# Number responders = 117
# max abundance in lab data = 60%
# mean abundance, dry burns = 0.15% +/- 14%

df.frac.of.total <- df.melt %>% subset(select = c(Sample, horizon, burn.trtmt, Veg.type, Abundance, DNA.type, incub.trtmt, Thermo.mid.max, site)) %>% group_by(Sample) %>% mutate(total.Abun = sum(Abundance))

p_FractionTotal <- ggplot(subset(df.frac.of.total, DNA.type == 'cDNA'), aes(x=site, y=total.Abun, fill=burn.trtmt)) +
  geom_bar(stat='identity',  position='dodge') + 
  facet_wrap(~horizon+incub.trtmt, labeller = labeller(horizon = HorizonLabels)) + 
  theme_bw()+
  labs(x= 'Sampling site ID',
     #  y = 'Decay rate constant for fast C pool \nlog(k1)',
       y = 'Fraction of total reads',
       fill = 'Burn treatment') + 
  scale_fill_manual(values = c('grey50','orange','red'),
                     limits = c('control','wet','dry'),
                     labels = c('Unburned\ncontrol','Moist soil\nburn','Dry soil\nburn')) +
  theme(axis.text.y = element_text(size = 14),
        legend.title = element_text(size=14),
        axis.text.x = element_text(size = 14),
        axis.title.x = element_text(size = 14, margin = margin(t=10, b=10)),
        axis.title.y = element_text(size = 14, margin = margin(r=10, l=10)),
        strip.text = element_text(size=14),
        legend.text = element_text(size=14),
        legend.position = '',
       # legend.background = element_rect(fill = "darkgray"),
        legend.box = "vertical")

df.frac.of.total$burn.trtmt <- factor(df.frac.of.total$burn.trtmt, levels = c('control','wet','dry'))

p_FractionTotal <- ggplot(subset(df.frac.of.total,incub.trtmt!='LIn'), aes(x=horizon, y=total.Abun, fill=burn.trtmt)) +
  geom_boxplot()+
  facet_wrap(~incub.trtmt+DNA.type, labeller = labeller(incub.trtmt = IncubLabels, DNA.type = DNALabels)) + 
  theme_bw()+
  labs(x= 'Soil horizon',
     #  y = 'Decay rate constant for fast C pool \nlog(k1)',
       y = 'Fraction of total reads',
       fill = 'Burn treatment') + 
  scale_fill_manual(values = c('grey50','orange','red'),
                     limits = c('control','wet','dry'),
                     labels = c('Unburned\ncontrol','Moist soil\nburn','Dry soil\nburn')) +
    scale_x_discrete(limits = c('O','A'),
                     labels = c('Organic','Mineral')) +
  theme(axis.text.y = element_text(size = 14),
        legend.title = element_text(size=14),
        axis.text.x = element_text(size = 14),
        axis.title.x = element_text(size = 14, margin = margin(t=10, b=10)),
        axis.title.y = element_text(size = 14, margin = margin(r=10, l=10)),
        strip.text = element_text(size=14),
        legend.text = element_text(size=14),
        legend.position = '',
       # legend.background = element_rect(fill = "darkgray"),
        legend.box = "vertical")

# Table summarize composition of responders by phylum:
df.phylum <- subset(df, select = c(mu.survival.max, mu.fast.max, mu.affinity.max, Phylum)) %>% unique()

df.phylum %>% mutate(Phylum = as.factor(Phylum)) %>% subset(mu.affinity.max>0) %>%subset(select = c(Phylum)) %>% View()

df.Class <- subset(df, select = c(mu.survival.max, mu.fast.max, mu.affinity.max, Class)) %>% unique()

df.Class %>% mutate(Class = as.factor(Class)) %>% subset(mu.fast.max>0) %>%subset(select = c(Class)) %>% summary()

# Count number of taxa by sample:
# Create a factor corresponding to the Genera
genfac = factor(tax_table(ps.norm.cDNA)[, "Genus"])
# Tabulate the counts for each genera in each sample
gentab = apply(otu_table(ps.norm.cDNA), MARGIN = 2, function(x) {
    tapply(x, INDEX = genfac, FUN = sum, na.rm = TRUE, simplify = TRUE)
})
head(gentab)[, 1:10]

x <- as.data.frame(gentab)
```

# New figures: 
```{r}
### BC dissim of burned vs. unburned on the x axis with a respiration metric on the y axis

# Import BC dissimilarity data
df.distances <- read.csv('../../data/bray-curtis-dissimilarites.csv')

# Combine this with metadata
df.distances <- merge(df.distances, subset(df.meta, burn.trtmt != 'control'))

# Organize
df.distances$incub.trtmt <- factor(df.distances$incub.trtmt, levels = c('pb','SI','LIwA','LIn'))
df.distances$horizon <- factor(df.distances$horizon, levels = c('O','A'))

# Import respiration data:
df.resp <- read.csv("../../data/fractional-c-remaining.csv")

df.merge <- merge(df.distances, df.resp, by = c('core.id','incub.trtmt')) %>%
  subset(incub.trtmt == 'SI') %>%
  subset(DayFactor == 34) %>%
  mutate(mass.C.respired = mass.C.remaining.g * (1-fractional.C.remaining))

pBC = ggplot(df.merge, aes(x=BC.distance.btw.burned.and.control, y=k1, color = horizon)) + 
  geom_point() + 
  scale_color_manual(values = c('darkgreen','grey70'))+
  stat_smooth(method=lm,se = FALSE,color='black') +
  theme_bw() + 
  #facet_wrap(~horizon, ncol=2,scales = 'fixed',
   #          labeller = labeller(horizon = HorizonLabels))+
  labs(y = 'Decay rate constant for \nfast C pool (k1)',
       x = 'Bray-Curtis dissimilarity \ncompared to unburned control')

summary(lm(data=df.merge,k1~BC.distance.btw.burned.and.control))

### Ordination colored by respiration metric:
# Full ordination
ps.norm.SI <- prune_samples(sample_data(ps.norm.full)$incub.trtmt =='SI', ps.norm.full)
ps.norm.SI <- prune_taxa(taxa_sums(ps.norm.SI)>0, ps.norm.SI)

ord.nmds = ordinate(ps.norm.SI,method = "NMDS",ddistance="wunifrac",k=3, weighted=TRUE)

# quick look at results:
plot_ordination(ps.norm.SI,ord.nmds,color="burn.trtmt",shape="incub.trtmt")
colnames(sample_data(ps.norm))

# Ordination plot with horizons and deg.hrs
p1 = plot_ordination(ps.norm.SI, ord.nmds, axes = c(1,2), color = "",
                      shape = 'burn.trtmt') + 
  #facet_grid(~horizon, labeller = labeller(horizon = HorizonLabels))+
  geom_point(size=3, alpha=0.8) +
  # scale_color_gradientn(colours = c('gold2','red','red4','black')) +
  # scale_color_manual(values = c('black','orange','red'),
  #                    labels = c('Unburned control', 'Wet soil burn','Wry soil burn'),
  #                    breaks = c('control','wet','dry'))+
  scale_shape_manual(values = c('circle','square','triangle'),
                     labels = c('Unburned control', 'Moist soil','Dry soil'),
                     breaks = c('control','wet','dry')) +
  #scale_shape_discrete(labels = c('Picea' = 'Picea spp.', 'Populus_tremuloides' = 'Populus tremuloides','Pinus_banksiana' = 'Pinus banksiana')) +
  #scale_shape_manual(values = c('circle','triangle'),
   #                  labels = c('O horizon','Mineral'),
    #                 breaks = c('O','A')) + 
  theme(legend.text = element_text(size = 12),
        axis.title = element_text(size=12),
        axis.text = element_text(size=12),
        legend.title = element_text(size=12)) + 
  labs(shape = 'Burn treatment',
       color = 'Fractional size\nof fast C pool') +
  theme_bw()

p1

## What fraction of total reads in dry soil burns 24 hr post burn were bacillus?
#  Bacillus fire surviving taxa = 8f0e2646c1389192d5a822f62bd295cf, 8adb5559e135233565acfea1287280bf
df <- read.csv('../../data/sequence-data/LibCombined/corncob-output/Manuscript/Trait-responder-OTUs.csv')

df.bac <- read.csv('../../data/sequence-data/LibCombined/corncob-output/Manuscript/Trait-responder-OTUs.csv') %>%
  subset(mu.survival.max>0) %>%
  subset(Genus=='Bacillus')

ps.norm.bac <- prune_samples(sample_data(ps.norm.gDNA)$incub.trtmt == 'pb', ps.norm.gDNA) 
ps.norm.bac <- prune_taxa(taxa_names(ps.norm.bac) %in% df.bac$OTU, ps.norm.bac)

bac <- psmelt(ps.norm.bac)
# Answer 9% (8f0e2646c1389192d5a822f62bd295cf) and 4% (8adb5559e135233565acfea1287280bf) of total reads in dry soil burns

### What percentage of all OTUs were identified as fast growers?
df.fast <- df %>%
  subset(mu.fast.max>0)

ps.norm.SI <- prune_samples(sample_data(ps.norm.gDNA)$incub.trtmt == 'SI', ps.norm.gDNA) 
ps.norm.SI <- prune_taxa(taxa_sums(ps.norm.SI)>0, ps.norm.SI)
ps.norm.SI
# Answer: IN the 5 week incubation, 0.4%

### Max relative abundance of responding taxa in lab samples?
ps.responders <- prune_taxa(taxa_names(ps.norm.gDNA) %in% df$OTU, ps.norm.gDNA)
df.responders <- psmelt(ps.responders)

df.responders <- df.responders %>%
  subset(select = c(Sample, Abundance, Full.id,burn.trtmt))

Num.responders <- df.responders %>% group_by(Full.id) %>% mutate(abun = sum(Abundance))
max(Num.responders$abun) # = 0.604

Num.responders.in.dry <- df.responders %>%
  subset(burn.trtmt == 'dry') %>% 
  group_by(Full.id) %>% 
  mutate(combined.abun = sum(Abundance))

mean(Num.responders.in.dry$combined.abun) # = 15
sd(Num.responders.in.dry$combined.abun) # =

```

# Figure 6: Trade offs
```{r}

df <- read.csv('../../data/sequence-data/LibCombined/corncob-output/Manuscript/all-responders.csv')
df.L2FC <- read.csv('../../data/sequence-data/LibCombined/corncob-output/Full_L2FC.csv')

# Checking to make sure that L2FC calculations didn't drop any OTUs. 
df.survival <- df %>%
  subset(experiment == 'survival'&
           T.cutoff %in% c('Low thermo > 50C','Mid thermo > 50C') & 
           intercept > 0) %>%
  subset(intercept > (mean(intercept)-1*sd(intercept)) ) %>% 
  subset(select = c(OTU,Phylum, Class, Order, Family, Genus, Species)) %>%
  unique()

df.survival.L2FC <- df.L2FC %>%
  subset(experiment == 'survival'  &
           T.cutoff %in% c('low thermo > 50C','Mid thermo > 50C') &
           mu > 0) %>%
  subset(mu > (mean(mu)-1*sd(mu))) %>%
  subset(select = c(OTU)) %>%
  unique()

df.survival.merge <- merge(df.survival, df.survival.L2FC)
#write.csv(df.survival,'../tables/survivors.csv', row.names=FALSE)

dim(df.survival)
dim(df.survival.L2FC)

df.fast <- df %>%
  subset(experiment == 'fast growth'&
           intercept > 0) %>%
  subset(intercept > (mean(intercept)-1*sd(intercept)) ) %>% 
  subset(select = c(OTU,Phylum, Class, Order, Family, Genus, Species)) %>%
  unique()
dim(df.fast)

df.fast.L2FC  <- df.L2FC %>%
  subset(experiment == 'fast growth'  &
           mu > 0) %>%
  subset(mu > (mean(mu)-1*sd(mu))) %>%
  subset(select = c(OTU)) %>%
  unique() 
dim(df.fast.L2FC)
df.fast.merge <- merge(df.fast.L2FC, df.fast)

df.affinity <- df %>%
  subset(experiment == 'affinity'&
           T.cutoff %in% c('Low thermo > 50','Mid thermo > 50') &
           intercept > 0) %>%
  subset(intercept > (mean(intercept)-1*sd(intercept)) ) %>% 
  subset(select = c(OTU,Phylum, Class, Order, Family, Genus, Species)) %>%
  unique()
dim(df.affinity)

df.affinity.L2FC  <- df.L2FC %>%
  subset(experiment == 'affinity'  &
           T.cutoff %in% c('low thermo > 50 C','mid thermo > 50 C') &
           mu > 0) %>%
  subset(mu > (mean(mu)-1*sd(mu))) %>%
  subset(select = c(OTU)) %>%
  unique() 
dim(df.affinity.L2FC)

df.affinity.merge <- merge(df.affinity.L2FC, df.affinity)

df.full <- df.L2FC

# Filter u ----
# First pull out enriched taxa:
df <-  subset(df.full, mu > 0)

# Create a dataframe for survivors filtering by temperature cutof and 1 sd from 
#   the mean of the intercept. Remove duplicate taxa. 
df.survival <- subset(df, experiment == 'survival' &
                        T.cutoff %in% c('low thermo > 50 C','mid thermo > 50 C')) %>%
  subset(mu > (mean(mu)-1*sd(mu))) %>%
  mutate(survival.L2FC = L2FC) %>%
  dplyr::group_by(OTU) %>%
  dplyr::mutate(max.mu = max(mu)) %>%
  subset(mu == max.mu) %>%
  subset(select = c(OTU, survival.L2FC)) %>%
  unique()%>%
  mutate(survivor = 'yes')

# Create dataframe for fast growers filtering for 1 sd from the mean: -----
df.fast <- subset(df, experiment == 'fast growth') %>%
  subset(mu > (mean(mu)-1*sd(mu))) %>%
  mutate(fast.L2FC = L2FC)%>%
  dplyr::group_by(OTU) %>%
  dplyr::mutate(max.mu = max(mu)) %>%
  subset(mu == max.mu) %>%
  subset(select = c(OTU, fast.L2FC)) %>%
  unique() %>%
  mutate(fast = 'yes')

# Create data frame for affinity taxa filtering for temperature cutoff and 1
#   sd from the mean:
df.affinity <- subset(df, experiment == 'affinity' &
                        T.cutoff %in% c('Low thermo > 50C','Mid thermo > 50C')) %>%
  subset(mu > (mean(mu)-1*sd(mu))) %>%
  mutate(affinity.L2FC = L2FC)%>%
  dplyr::group_by(OTU) %>%
  dplyr::mutate(max.mu = max(mu)) %>%
  subset(mu == max.mu) %>%
  subset(select = c(OTU, affinity.L2FC)) %>%
  unique()%>%
  mutate(affinity = 'yes')

# Combine the three dataframes:
df.index <- merge(df.survival, df.fast, all = TRUE) %>%
  merge(df.affinity, all = TRUE) %>%
  subset(select = c(OTU, survivor, fast, affinity))

df.ids <- read.csv('../data/sequence-data/LibCombined/corncob-output/all-responders.csv') %>%
  subset(select = c(OTU, Domain, Phylum, Class, Order, Family, Genus, Species)) %>%
  unique()
df.ids <- merge(df.index,df.ids, all = FALSE)
write.csv(df.ids, '../tables/all-responding-otus.csv', row.names = FALSE)

# Next take the full responder dataset and filter out any duplicate taxa

x.1 <- df.full %>%
  subset(experiment == 'survival') %>%
  subset(select = c(OTU, mu, experiment,horizon,
                    fold.change, L2FC)) %>%
  dplyr::group_by(OTU, experiment) %>%
  dplyr::mutate(max.mu = max(abs(mu))) %>%
  subset(max.mu == abs(mu)) %>%
  dplyr::mutate(survival.L2FC = L2FC) %>%
  subset(select = c(OTU, survival.L2FC)) %>%
  unique()

x.2 <- df.full %>%
  subset(experiment == 'fast growth') %>%
  subset(select = c(OTU, mu, experiment,horizon,
                    fold.change, L2FC)) %>%
  dplyr::group_by(OTU, experiment) %>%
  dplyr::mutate(max.mu = max(abs(mu))) %>%
  subset(max.mu == abs(mu)) %>%
  dplyr::mutate(fast.L2FC = L2FC) %>%
  subset(select = c(OTU, fast.L2FC)) %>%
  unique()

x.3 <- df.full %>%
  subset(experiment == 'affinity') %>%
  subset(select = c(OTU, mu, experiment,horizon,
                    fold.change, L2FC)) %>%
  dplyr::group_by(OTU, experiment) %>%
  dplyr::mutate(max.mu = max(abs(mu))) %>%
  subset(max.mu == abs(mu)) %>%
  dplyr::mutate(affinity.L2FC = L2FC) %>%
  subset(select = c(OTU, affinity.L2FC)) %>%
  unique()

x <- merge(x.1, x.2, all = TRUE) %>%
  merge(x.3, all = TRUE)

# Create list of responding taxa:
responding.OTU.list = unique(df.index$OTU)

# Then filter this to include only taxa identified as responders (i.e. meeting
#   the respective cutoffs for at least one of the three trait experiments)
df.traits <- filter(x, OTU %in% responding.OTU.list) %>% 
  unique()

df.traits <- merge(df.index, df.traits)

#df.traits <- test %>%
#  merge(unique(subset(df, select = c(OTU, Domain, Phylum, Class, Order, Family, Genus, Species))))

df.traits$trait = 'null'

for (i in 1:nrow(df.traits)) {
  if (df.traits$survivor[i] == 'yes' &
      df.traits$survival.L2FC[i] > 0 &
      is.na(df.traits$fast[i]) == TRUE &
      is.na(df.traits$affinity[i]) == TRUE) {
    df.traits$trait[i] = 'survival'
      } else if (df.traits$fast[i] == 'yes' &
                 df.traits$fast.L2FC[i] > 0 &
                 is.na(df.traits$survivor[i]) == TRUE &
                 is.na(df.traits$affinity[i]) == TRUE) {
        df.traits$trait[i] = 'fast growth'
      } else if (df.traits$affinity[i] == 'yes' &
                 df.traits$affinity.L2FC[i] > 0 &
                 is.na(df.traits$survivor[i]) == TRUE &
                 is.na(df.traits$fast[i]) == TRUE) {
        df.traits$trait[i] = 'affinity'
      } else if (df.traits$survivor[i] == 'yes' &
                 df.traits$survival.L2FC[i] > 0 &
                 df.traits$fast[i] == 'yes' &
                 df.traits$fast.L2FC[i] > 0 &
                 is.na(df.traits$affinity[i]) == TRUE) {
        df.traits$trait[i] = 'survival and fast growth'
      } else if (df.traits$fast[i] == 'yes' &
                 df.traits$fast.L2FC[i] > 0 &
                 df.traits$affinity[i] == 'yes' &
                 df.traits$affinity.L2FC[i] > 0 &
                 is.na(df.traits$survivor[i]) == TRUE) {
        df.traits$trait[i] = 'fast growth and affinity'
      } else if (df.traits$survivor[i] == 'yes' &
                 df.traits$survival.L2FC[i] > 0 &
                 df.traits$affinity[i] == 'yes' &
                 df.traits$affinity.L2FC[i] > 0 &
                 is.na(df.traits$fast[i]) == TRUE) {
        df.traits$trait[i] = 'survival and affinity'
      } else if (df.traits$survivor[i] == 'yes' &
                 df.traits$survival.L2FC[i] > 0 &
                 df.traits$fast[i] == 'yes' &
                 df.traits$fast.L2FC[i] > 0 &
                 df.traits$affinity[i] == 'yes' &
                 df.traits$affinity.L2FC[i] > 0) {
        df.traits$trait[i] = 'survival, fast growth, \nand affinity'
      } else {
        df.traits$trait[i] = 'none'
      } 
}

unique(df.traits$trait)

head(df.traits)

library(RColorBrewer)
old.trait.colors = c('gold','green','royalblue3','purple','darkred','orange','black')
trait.colors = c(brewer.pal(n=10, name = 'Spectral'), 'black')

trait.colors = c('#E6AB02','#33A02C','#386CB0','#762A83','#D73027','#F46D43','black')

# Finally build trait trade-off plots:
p1 = ggplot(df.traits) + 
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 0)+ 
  geom_point(aes(x=survival.L2FC, y=fast.L2FC, color = trait, fill = trait), shape = 21,size=4, alpha=1) + 
  theme_bw() +
  labs(x= expression(Survival~(log[2]-fold~change)),
       #  y = 'Decay rate constant for fast C pool \nlog(k1)',
       y = expression(Fast~growth~(log[2]-fold~change))) + 
  scale_color_manual(limits = c('survival','survival and fast growth','fast growth','fast growth and affinity',
                                'affinity','affinity and survival','survival, fast growth, \nand affinity'),
                     values = trait.colors) + 
  scale_fill_manual(limits = c('survival','survival and fast growth','fast growth','fast growth and affinity',
                                'affinity','affinity and survival','survival, fast growth, \nand affinity'),
                     values = trait.colors) + 
  theme(axis.text.y = element_text(size = 14),
        legend.title = element_text(size=10),
        axis.text.x = element_text(size = 14),
        axis.title.x = element_text(size = 14, margin = margin(t=10, b=10)),
        axis.title.y = element_text(size = 14, margin = margin(r=10, l=10)),
        strip.text = element_text(size=14),
        legend.text = element_text(size=10),
        legend.position = 'none',
        # legend.background = element_rect(fill = "darkgray"),
        legend.box = "vertical")+
  stat_smooth(aes(x=survival.L2FC, y=fast.L2FC),method=lm, se=FALSE, color = 'grey40',
              geom = 'smooth')

p2 = ggplot(df.traits) + 
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 0)+ 
  geom_point(aes(x=fast.L2FC, y=affinity.L2FC, color = trait, fill=trait), shape = 21,  size=4,alpha=1) + 
  theme_bw() +
  labs(x= expression(Fast~growth~(log[2]-fold~change)),
       #  y = 'Decay rate constant for fast C pool \nlog(k1)',
       y = expression(Post~fire~envir.~affinity~(log[2]-fold~change))) + 
  scale_color_manual(limits = c('survival','survival and fast growth','fast growth','fast growth and affinity',
                                'affinity','affinity and survival','survival, fast growth, \nand affinity'),
                     values = trait.colors) + 
  scale_fill_manual(limits = c('survival','survival and fast growth','fast growth','fast growth and affinity',
                                'affinity','affinity and survival','survival, fast growth, \nand affinity'),
                     values = trait.colors) + 
  theme(axis.text.y = element_text(size = 14),
        legend.title = element_text(size=10),
        axis.text.x = element_text(size = 14),
        axis.title.x = element_text(size = 14, margin = margin(t=10, b=10)),
        axis.title.y = element_text(size = 14, margin = margin(r=10, l=10)),
        strip.text = element_text(size=14),
        legend.text = element_text(size=10),
        legend.position = 'none',
        # legend.background = element_rect(fill = "darkgray"),
        legend.box = "vertical")

p3 = ggplot(df.traits) + 
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 0)+ 
  geom_point(aes(x=affinity.L2FC, y=survival.L2FC, color = trait, fill = trait), 
             shape = 21,size=4,alpha=1) + 
  theme_bw() +
  labs(x=expression(Post~fire~envir.~affinity~(log[2]-fold~change)),
       #  y = 'Decay rate constant for fast C pool \nlog(k1)',
       y = expression(Survival~(log[2]-fold~change)),
       color = 'Bacterial strategy:',
       fill = 'Bacterial strategy:') + 
  scale_color_manual(limits = c('survival','survival and fast growth','fast growth','fast growth and affinity',
                                'affinity','affinity and survival','survival, fast growth, \nand affinity'),
                     values = trait.colors) + 
  scale_fill_manual(limits = c('survival','survival and fast growth','fast growth','fast growth and affinity',
                                'affinity','affinity and survival','survival, fast growth, \nand affinity'),
                     values = trait.colors) + 
  theme(axis.text.y = element_text(size = 14),
        legend.title = element_text(size=14),
        axis.text.x = element_text(size = 14),
        axis.title.x = element_text(size = 14, margin = margin(t=10, b=10)),
        axis.title.y = element_text(size = 14, margin = margin(r=10, l=10)),
        strip.text = element_text(size=14),
        legend.text = element_text(size=14),
        legend.position = 'none',
        # legend.background = element_rect(fill = "darkgray"),
        legend.box = "vertical")
cowplot::plot_grid(p1, p2,p3, ncol=3)

test <- lm(fast.L2FC ~ affinity.L2FC, data = df.traits)
summary(test)
test
#

x <- grDevices::rainbow(n=6)

df <- read.csv('../data/sequence-data/LibCombined/corncob-output/all-responders.csv')
df <- subset(df, select = c(OTU, Domain, Phylum, Class, Order, Family, Genus, Species)) %>%
  unique()

df.traits <- merge(df.traits,df)
dim(df.traits)
```

```{r}
# How much of total community do affinity taxa account for?
ps.norm.LI <- prune_samples(sample_data(ps.norm.gDNA)$incub.trtmt == 'LIwA', ps.norm.gDNA)
ps.norm.LI <- prune_taxa(taxa_names(ps.norm.LI) %in% df.affinity.L2FC$OTU, ps.norm.LI)

df.norm.LI <- psmelt(ps.norm.LI)
df.norm.LI$burn.trtmt = factor(df.norm.LI$burn.trtmt, levels = c('control','wet','dry'))

df.norm.LI <- df.norm.LI %>%
  group_by(core.id.hor.incub, horizon, burn.trtmt) %>%
  summarize(sum.Abundance = sum(Abundance))

p = ggplot(df.norm.LI, aes(x=burn.trtmt, y = sum.Abundance, color = horizon)) + 
  geom_point() + 
  facet_grid(~horizon) + 
  theme_bw()
```

# PERMANOVAs:
```{r}
ps.E1 <- prune_samples(sample_data(ps.norm.full)$incub.trtmt == 'pb' &
                         sample_data(ps.norm.full)$DNA.type == 'cDNA', ps.norm.full)
ps.E1 <- prune_taxa(taxa_sums(ps.E1)>0, ps.E1)
ps.E1

ps.E1.DNA <- prune_samples(sample_data(ps.norm.full)$incub.trtmt == 'pb' &
                         sample_data(ps.norm.full)$DNA.type == 'gDNA', ps.norm.full)
ps.E1.DNA <- prune_taxa(taxa_sums(ps.E1.DNA)>0, ps.E1.DNA)
ps.E1.DNA

ps.E2 <- prune_samples(sample_data(ps.norm.full)$incub.trtmt == 'SI' &
                         sample_data(ps.norm.full)$DNA.type == 'gDNA', ps.norm.full)
ps.E2 <- prune_taxa(taxa_sums(ps.E2)>0, ps.E2)
ps.E2

ps.E3  <- prune_samples(sample_data(ps.norm.full)$incub.trtmt == 'LIwA' &
                         sample_data(ps.norm.full)$DNA.type == 'gDNA', ps.norm.full)
ps.E3 <- prune_taxa(taxa_sums(ps.E3)>0, ps.E3)
ps.E3

MyWunifrac_E1 <- distance(ps.E1, method = 'wunifrac')
MyWunifrac_E1.DNA <- distance(ps.E1.DNA, method = 'wunifrac')
MyWunifrac_E2 <- distance(ps.E2, method = 'wunifrac')
MyWunifrac_E3 <- distance(ps.E3, method = 'wunifrac')
MyWunifrac_All <- distance(ps.norm.gDNA, method = 'wunifrac')

SamDat_E1 = data.frame(sample_data(ps.E1))
SamDat_E1.DNA = data.frame(sample_data(ps.E1.DNA))
SamDat_E2 = data.frame(sample_data(ps.E2))
SamDat_E3 = data.frame(sample_data(ps.E3))
SamDat_All = data.frame(sample_data(ps.norm.gDNA))

test_E1 <- adonis(MyWunifrac_E1 ~ Veg.type+PRE.hor.thickness.cm+pH+percent.total.C+
                    percent.total.N+texture+burn.trtmt+horizon, data = SamDat_E1)
test_E1.DNA <- adonis(MyWunifrac_E1.DNA ~ Veg.type+PRE.hor.thickness.cm+pH+percent.total.C+
                    percent.total.N+texture+burn.trtmt+horizon, data = SamDat_E1.DNA)
test_E2 <- adonis(MyWunifrac_E2 ~ Veg.type+PRE.hor.thickness.cm+pH+percent.total.C+
                    percent.total.N+texture+burn.trtmt+horizon, data = SamDat_E2)
test_E3 <- adonis(MyWunifrac_E3 ~ Veg.type+PRE.hor.thickness.cm+pH+percent.total.C+
                    percent.total.N+texture+burn.trtmt+horizon, data = SamDat_E3)
test_All <- adonis(MyWunifrac_All ~ incub.trtmt+burn.trtmt+pH+horizon+Veg.type, data = SamDat_All)

test_E1
test_E1.DNA
test_E2
test_E3
test_All

test_E1_sub <- adonis(MyWunifrac_E1.DNA ~ percent.total.N, data = SamDat_E1.DNA)
test_E1_sub

# Fire survival
adonis(MyWunifrac_E1 ~ Veg.type, data = SamDat_E1)
adonis(MyWunifrac_E1 ~ PRE.hor.thickness.cm, data = SamDat_E1)
adonis(MyWunifrac_E1 ~ pH, data = SamDat_E1)
adonis(MyWunifrac_E1 ~ percent.total.C, data = SamDat_E1)
adonis(MyWunifrac_E1 ~ percent.total.N, data = SamDat_E1)
adonis(MyWunifrac_E1 ~ texture, data = SamDat_E1)
adonis(MyWunifrac_E1 ~ burn.trtmt, data = SamDat_E1)
adonis(MyWunifrac_E1 ~ horizon, data = SamDat_E1)

adonis(MyWunifrac_E1.DNA ~ Veg.type, data = SamDat_E1.DNA)
adonis(MyWunifrac_E1.DNA ~ PRE.hor.thickness.cm, data = SamDat_E1.DNA)
adonis(MyWunifrac_E1.DNA ~ pH, data = SamDat_E1.DNA)
adonis(MyWunifrac_E1.DNA ~ percent.total.C, data = SamDat_E1.DNA)
adonis(MyWunifrac_E1.DNA ~ percent.total.N, data = SamDat_E1.DNA)
adonis(MyWunifrac_E1.DNA ~ texture, data = SamDat_E1.DNA)
adonis(MyWunifrac_E1.DNA ~ burn.trtmt, data = SamDat_E1.DNA)
adonis(MyWunifrac_E1.DNA ~ horizon, data = SamDat_E1.DNA)

# Fast growth
adonis(MyWunifrac_E2 ~ Veg.type, data = SamDat_E2)
adonis(MyWunifrac_E2 ~ PRE.hor.thickness.cm, data = SamDat_E2)
adonis(MyWunifrac_E2 ~ pH, data = SamDat_E2)
adonis(MyWunifrac_E2 ~ percent.total.C, data = SamDat_E2)
adonis(MyWunifrac_E2 ~ percent.total.N, data = SamDat_E2)
adonis(MyWunifrac_E2 ~ texture, data = SamDat_E2)
adonis(MyWunifrac_E2 ~ burn.trtmt, data = SamDat_E2)
adonis(MyWunifrac_E2 ~ horizon, data = SamDat_E2)

# Test of autoclaving
ps.E3.autoclave  <- prune_samples(sample_data(ps.norm.full)$incub.trtmt %in% c('LIwA','LIn'), ps.norm.full)
ps.E3.autoclave <- prune_taxa(taxa_sums(ps.E3.autoclave)>0, ps.E3.autoclave)
ps.E3.autoclave
SamDat <- data.frame(sample_data(ps.E3.autoclave))
head(SamDat,5)

MyWunifrac_E3.autoclave <- distance(ps.E3.autoclave, method = 'wunifrac')

SamDat_E3.autoclaving = data.frame(sample_data(ps.E3.autoclave))

adonis(MyWunifrac_E3.autoclave ~ Veg.type+PRE.hor.thickness.cm+pH+percent.total.C+
                    percent.total.N+texture+burn.trtmt+horizon + incub.trtmt, data = SamDat_E3.autoclaving)

adonis(MyWunifrac_E3.autoclave ~ Veg.type, data = SamDat_E3.autoclaving)
adonis(MyWunifrac_E3.autoclave ~ PRE.hor.thickness.cm, data = SamDat_E3.autoclaving)
adonis(MyWunifrac_E3.autoclave ~ pH, data = SamDat_E3.autoclaving)
adonis(MyWunifrac_E3.autoclave ~ percent.total.C, data = SamDat_E3.autoclaving)
adonis(MyWunifrac_E3.autoclave ~ percent.total.N, data = SamDat_E3.autoclaving)
adonis(MyWunifrac_E3.autoclave ~ texture, data = SamDat_E3.autoclaving)
adonis(MyWunifrac_E3.autoclave ~ burn.trtmt, data = SamDat_E3.autoclaving)
adonis(MyWunifrac_E3.autoclave ~ horizon, data = SamDat_E3.autoclaving)
adonis(MyWunifrac_E3.autoclave ~ incub.trtmt, data = SamDat_E3.autoclaving)

# All
adonis(MyWunifrac_All ~ Veg.type, data = SamDat_All)
adonis(MyWunifrac_All ~ pH, data = SamDat_All)
adonis(MyWunifrac_All ~ incub.trtmt, data = SamDat_All)
adonis(MyWunifrac_All ~ burn.trtmt, data = SamDat_All)
adonis(MyWunifrac_All ~ horizon, data = SamDat_All)
```

# Figure 7:
Ordinations
```{r}
# Full ordination
ps.full.dataset <- readRDS('../../CombinedFireSim1519/ps.1519FireSim.merged')
tree <- ape::read.tree('../../CombinedFireSim1519/tree/merged-exported-tree/tree.nwk')

tree <- phy_tree(tree)
tax <- tax_table(ps.full.dataset)
otu <- otu_table(ps.full.dataset)
sam <- sample_data(ps.full.dataset)

ps.full.dataset = phyloseq::phyloseq(otu,tax, sam,tree)
ps.full.dataset

# Setting some plot parameters
year.labs = c("One year post-fire","Five years post-fire")
names(year.labs) = c(1,5)
horizon.labs = c("Mineral","Organic")

# Playing with new datasets
ps = ps.full.dataset

# Add column with pairs info
sample_data(ps)$Pairs = paste(sample_data(ps)$Site_ID,sample_data(ps)$Org_or_Min,sep="_")
head(sample_data(ps)$Pairs)

# Normalize sample counts
ps.norm = transform_sample_counts(ps,function(x) x/sum(x))
sample_data(ps.norm)$Study = sample_data(ps.norm)$Years_Since_Fire
sample_data(ps.norm)$Study[is.na(sample_data(ps.norm)$Study)]="Lab"

ord.nmds = ordinate(ps.norm,method = "NMDS",ddistance="wunifrac",k=3, weighted=TRUE)

plot_ordination(ps.norm,ord.pcoa,color="burn.trtmt",shape="incub.trtmt")
plot_ordination(ps.norm,ord.nmds,color="pH",shape="Study")
# Neat, they all fall out on top of each other. That's nice.
colnames(sample_data(ps.norm))

ps.norm.gDNA <- prune_samples(sample_data(ps.norm.gDNA)$incub.trtmt %in% c('pb','SI','LIwA'), ps.norm.gDNA)
ps.norm.gDNA <- prune_taxa(taxa_sums(ps.norm.gDNA)>0,ps.norm.gDNA)

MyNMDS.full.gDNA <- ordinate(ps.norm.gDNA, method = 'NMDS', k=3, ddistance="wunifrac", weighted=TRUE)

# Ordination plot with horizons and deg.hrs
p1 = plot_ordination(ps.norm.gDNA, MyNMDS.full.gDNA, axes = c(1,2), 
                      color = "burn.trtmt",
                      shape = 'incub.trtmt') + 
  facet_grid(~horizon, labeller = labeller(horizon = HorizonLabels,
                                                   DNA.type = DNALabels,
                                                   incub.trtmt = IncubLabels))+
  #facet_grid(~horizon, labeller = labeller(horizon = HorizonLabels))+
  geom_point(size=3, alpha=0.8) +
  #geom_line(aes(group = core.id.hor)) +
  labs(color = 'Burn treatment',
       #color = expression("Degree ("*degree*C*") hours"), 
       shape = 'experiment') +
  #scale_color_gradientn(colours = c('gold2','red','red4','black')) +
  scale_color_manual(values = c('black','orange','red'),
                     labels = c('Unburned control', 'Wet soil burn','Wry soil burn'),
                     breaks = c('control','wet','dry'))+
  scale_shape_manual(values = c('circle','square','triangle'),
                     labels = c('Fire survival', 'Fast growth','Post-fire envir. affinity'),
                     breaks = c('pb','SI','LIwA')) +
  #scale_shape_discrete(labels = c('Picea' = 'Picea spp.', 'Populus_tremuloides' = 'Populus tremuloides','Pinus_banksiana' = 'Pinus banksiana')) +
  #scale_shape_manual(values = c('circle','triangle'),
   #                  labels = c('O horizon','Mineral'),
    #                 breaks = c('O','A')) + 
  theme(legend.text = element_text(face = 'italic',size = 16),
        axis.title = element_text(size=16),
        axis.text = element_text(size=16),
        legend.title = element_text(size=16)) + 
  theme_bw()

p1
```

# Supplementary
```{r}
# Carbon and nitrogen loss
pChange_C <- ggplot(subset(df.meta, incub.trtmt == 'SI')) + 
  geom_boxplot(aes(x=burn.trtmt, y = percent.total.C, fill = horizon), alpha = 0.7, varwidth = FALSE) + 
    theme_bw()+
  labs(x= 'Burn treatment',
     #  y = 'Decay rate constant for fast C pool \nlog(k1)',
       y = 'Total C (%)',
       fill = 'Soil horizon') + 
  scale_fill_manual(values = c('black','grey90'),
                     limits = c('O','A'),
                     labels = c('Organic','Mineral')) +
  scale_x_discrete(limits = c('control','wet','dry'),
                     labels = c('Unburned\ncontrol','Wet soil\nburn','Dry soil\nburn')) +
  theme(axis.text.y = element_text(size = 14),
        legend.title = element_text(size=14),
        axis.text.x = element_text(size = 14),
        axis.title.x = element_text(size = 14, margin = margin(t=10, b=10)),
        axis.title.y = element_text(size = 14, margin = margin(r=10, l=10)),
        strip.text = element_text(size=14),
        legend.text = element_text(size=14),
        legend.position = '',
       # legend.background = element_rect(fill = "darkgray"),
        legend.box = "vertical")

pChange_N <- ggplot(subset(df.meta, incub.trtmt == 'SI')) + 
  geom_boxplot(aes(x=burn.trtmt, y = percent.total.N, fill = horizon), alpha = 0.7, varwidth = FALSE) + 
    theme_bw()+
  labs(x= 'Burn treatment',
     #  y = 'Decay rate constant for fast C pool \nlog(k1)',
       y = 'Total N (%)',
       fill = 'Soil horizon') + 
  scale_fill_manual(values = c('black','grey90'),
                     limits = c('O','A'),
                     labels = c('Organic','Mineral')) +
  scale_x_discrete(limits = c('control','wet','dry'),
                     labels = c('Unburned\ncontrol','Wet soil\nburn','Dry soil\nburn')) +
  theme(axis.text.y = element_text(size = 14),
        legend.title = element_text(size=14),
        axis.text.x = element_text(size = 14),
        axis.title.x = element_text(size = 14, margin = margin(t=10, b=10)),
        axis.title.y = element_text(size = 14, margin = margin(r=10, l=10)),
        strip.text = element_text(size=14),
        legend.text = element_text(size=14),
        legend.position = '',
       # legend.background = element_rect(fill = "darkgray"),
        legend.box = "vertical")

cowplot::plot_grid(pChange_C, pChange_N, ncol=2)

test <- aov(percent.total.N ~ burn.trtmt, data = subset(df.meta, incub.trtmt == 'SI' & horizon == 'O'))
summary(test)
TukeyHSD(test)

# Traits
df.resp.taxa <- merge(df.survival.L2FC, df.fast.L2FC, all = TRUE) %>% merge(df.affinity.L2FC, all = TRUE)

ps.taxa <- prune_taxa(taxa_names(ps.norm.full) %in% df.resp.taxa$OTU, ps.norm.full)
ps.taxa
df.taxa <- psmelt(ps.taxa)

# survival
ps.survival.taxa <- prune_taxa(taxa_names(ps.norm.full) %in% df.survival.L2FC$OTU, ps.norm.full)
ps.survival.taxa

df.survival.taxa <- psmelt(ps.survival.taxa)

df.survival.taxa$burn.trtmt <- factor(df.survival.taxa$burn.trtmt, levels = c('control','wet','dry'))

p = ggplot(subset(df.survival.taxa, incub.trtmt == 'pb' & DNA.type == 'gDNA'), aes(x=site, y = Abundance)) + 
  geom_bar(aes(fill = burn.trtmt),stat = 'identity',position = 'dodge') + 
  theme_bw()+
  facet_wrap(~horizon,ncol=1) + 
  scale_fill_manual(values = c('black','slategray3','tan3'),
                     limits = c('control','wet','dry'),
                     labels = c('control','wet soil','dry soil'))+
  labs(fill = 'Burn treatment',
       x = 'site',
       y='Total abundance (%)') + 
 theme(axis.text.y = element_text(size = 14),
        legend.title = element_text(size=14),
        axis.text.x = element_text(size = 14),
        axis.title.x = element_text(size = 14, margin = margin(t=10, b=10)),
        axis.title.y = element_text(size = 14, margin = margin(r=10, l=10)),
        strip.text = element_text(size=14),
        legend.text = element_text(size=14),
        legend.position = '',
       # legend.background = element_rect(fill = "darkgray"),
        legend.box = "vertical")

# fast
ps.fast.taxa <- prune_taxa(taxa_names(ps.norm.full) %in% df.fast.L2FC$OTU, ps.norm.full)
ps.fast.taxa

df.fast.taxa <- psmelt(ps.fast.taxa)

p = ggplot(subset(df.survival.taxa, incub.trtmt == 'SI' & DNA.type == 'gDNA'), aes(x=site, y = Abundance)) + 
  geom_bar(aes(fill = burn.trtmt),stat = 'identity',position = 'dodge') + 
  theme_bw()+
  facet_wrap(~horizon,ncol=1) + 
  scale_fill_manual(values = c('black','slategray3','tan3'),
                     limits = c('control','wet','dry'),
                     labels = c('control','wet soil','dry soil'))+
  labs(fill = 'Burn treatment',
       x = 'site',
       y='Total abundance (%)') + 
 theme(axis.text.y = element_text(size = 14),
        legend.title = element_text(size=14),
        axis.text.x = element_text(size = 14),
        axis.title.x = element_text(size = 14, margin = margin(t=10, b=10)),
        axis.title.y = element_text(size = 14, margin = margin(r=10, l=10)),
        strip.text = element_text(size=14),
        legend.text = element_text(size=14),
        legend.position = '',
       # legend.background = element_rect(fill = "darkgray"),
        legend.box = "vertical")

# affinity
ps.affinity.taxa <- prune_taxa(taxa_names(ps.norm.full) %in% df.affinity.L2FC$OTU, ps.norm.full)
ps.affinity.taxa

df.affinity.taxa <- psmelt(ps.affinity.taxa)

p = ggplot(subset(df.survival.taxa, incub.trtmt == 'LIwA' & DNA.type == 'gDNA'), aes(x=site, y = Abundance)) + 
  geom_bar(aes(fill = burn.trtmt),stat = 'identity',position = 'dodge') + 
  theme_bw()+
  facet_wrap(~horizon,ncol=1) + 
  scale_fill_manual(values = c('black','slategray3','tan3'),
                     limits = c('control','wet','dry'),
                     labels = c('control','wet soil','dry soil'))+
  labs(fill = 'Burn treatment',
       x = 'site',
       y='Total abundance (%)') + 
 theme(axis.text.y = element_text(size = 14),
        legend.title = element_text(size=14),
        axis.text.x = element_text(size = 14),
        axis.title.x = element_text(size = 14, margin = margin(t=10, b=10)),
        axis.title.y = element_text(size = 14, margin = margin(r=10, l=10)),
        strip.text = element_text(size=14),
        legend.text = element_text(size=14),
        legend.position = 'right',
       # legend.background = element_rect(fill = "darkgray"),
        legend.box = "vertical")

```

FIN --------------------------------

EXTRAS:

# SI: plot responders as fraction of total comm. vs. total C mineralized over 5 week incub.
```{r}
### Goal plot with fraction of total comm by trait category vs. total C-CO2 respired after 5 weeks

# 1. import trait data
df <- read.csv('../../data/sequence-data/LibCombined/corncob-output/Manuscript/Trait-responder-OTUs.csv')
head(df)

df$trait = 'TBD'

# assign traits:
for (i in 1:nrow(df)) {
  if (is.na(df$mu.fast.max[i]) & is.na(df$mu.affinity.max[i])) {
    df$trait[i] = 'Survivor' 
  } else if (is.na(df$mu.survival.max[i]) & is.na(df$mu.affinity.max[i])) {
    df$trait[i] = 'Fast grower' 
  } else if (is.na(df$mu.survival.max[i]) & is.na(df$mu.fast.max[i])) {
    df$trait[i] = 'Affinity taxa' 
  } else if (is.na(df$mu.survival.max[i])) {
    df$trait[i] = 'Fast grower and affinity taxa'
  } else if (is.na(df$mu.affinity.max[i])) {
    df$trait[i] = 'Survivor and fast grower'
  } else if (is.na(df$mu.fast.max[i])) {
    df$trait[i] = 'Survivor and affinity taxa' 
  } else {
    df$trait[i] = 'ALL'
  }
}

# 2. import full comm data at end of SI incubation
ps.norm.SI <- prune_samples(sample_data(ps.norm.full)$incub.trtmt == 'SI', ps.norm.full)

# 3. Assign traits to OTUs in the SI comm data
ps.norm.SI <- prune_taxa(taxa_names(ps.norm.SI) %in% df$OTU, ps.norm.SI)
ps.norm.SI <- prune_taxa(taxa_sums(ps.norm.SI)>0, ps.norm.SI)

df.norm.SI <- psmelt(ps.norm.SI)

df.norm.SI <- merge(df.norm.SI, subset(df, select = c(OTU, trait)))

p = ggplot(df.norm.SI, aes(x=burn.trtmt, y=Abundance)) + 
  geom_bar(aes(fill=trait), stat='identity', position='stack')

# 4. Import respiration data
#df.resp <- read.csv("../../data/fractional-c-remaining.csv")
# df.resp <- df (imported from "incubations_microbial_respiration_TLWcheck.R")

# 5. sum total respiration at end of 5 weeks
df.resp.SI <- subset(df.resp, Incubation.trtmt == 'SI')

# 6. merge trait and respiration dataframes
df.sum <- df.norm.SI %>%
  group_by(Full.id, trait) %>%
  mutate(trait.abun = sum(Abundance)) %>%
  subset(select = c(core.id,trait,trait.abun)) %>%
  unique() %>%
  merge(df.resp.SI, by = 'core.id')

# 7. plot fraction of total comm vs. total C-CO2 respired
p = ggplot(subset(df.sum, Day==34), aes(x=trait.abun, y = C.CO2.absorbed.per.g.C)) + 
  geom_point(aes(color=trait)) + 
  facet_wrap(~burn.trtmt)+
  theme_bw()

p = ggplot(df.sum, aes(x=(1-fractional.C.remaining),y=trait.abun)) + 
  geom_bar(aes(fill=trait), stat='identity', position='stack') + 
  theme_bw()

# Initial mineralization after 2-7 days

```

# figuring out corncob + trait assignments
```{r}
df <- read.csv('../data/sequence-data/LibCombined/corncob-output/all-responders.csv')

unique(subset(df, select = c(experiment, DNA.type, Comparison, pH.cutoff, Temp.cutoff)))

df.LI <- read.csv('../data/sequence-data/LibCombined/corncob-output/Ex3 Affinity/all-responders.csv')

df$horizon <- factor(df$horizon, levels = c('O','A'))
df$experiment <- factor(df$experiment, levels = c('survival','fast growth','affinity'))

for (i in 1:nrow(df)) {
  if (abs(df$intercept[i]) > 2) {
    df$trimmed.intercept[i] = df$intercept[i]
  } else {
    df$trimmed.intercept[i] = NA
  }
}

p = ggplot(df) + 
  geom_histogram(aes(intercept),fill = 'red', binwidth=1) + 
  #geom_histogram(aes(trimmed.intercept), fill = 'blue', binwidth=1) + 
  theme_bw() + 
  facet_grid(experiment+DNA.type~horizon, scales = 'free')

length(unique(subset(df, Temp.cutoff == 'Low thermo greater than 50 C' &
                       experiment == 'survival')$OTU))

p1 = ggplot(subset(df, experiment == 'survival'), aes(intercept)) +
  geom_histogram() + 
  theme_bw() + 
  facet_grid(horizon+Comparison~Temp.cutoff, scales = 'free')

p2 = ggplot(subset(df, experiment == 'fast growth'), aes(intercept)) +
  geom_histogram() + 
  theme_bw() + 
  facet_grid(horizon+Comparison~pH.cutoff, scales = 'free')

p3 = ggplot(subset(df, experiment == 'survival'), aes(intercept)) +
  geom_histogram() + 
  theme_bw() + 
  facet_grid(horizon+Comparison~Temp.cutoff, scales = 'free')

length(unique(subset(df, intercept > 0 & experiment == 'affinity')$OTU))
length(unique(subset(df, intercept < 0 & experiment == 'affinity')$OTU))
length(unique(subset(df, intercept > 1 & experiment == 'affinity')$OTU))
length(unique(subset(df, intercept < -1 & experiment == 'affinity')$OTU))

dim(subset(df, experiment == 'fast growth' &
             horizon == 'O' &
             Comparison == 'pb.dry.v.SI.dry' &
             pH.cutoff != 'null'))
             #Temp.cutoff != 'null'))

```

# PLotting rel. abundance of responders:
Reallying interesting data. We're seeing interesting trends with positive responders.
The relative abundance of positive responders increases across the dry soil post-burn
pH range, but decreases across the control soil pH range. 
```{r}
df.full.OTU <- read.csv('../data/sequence-data/LibCombined/corncob-output/all-resopnders.csv')
df.full3 <- read.csv('../data/sequence-data/LibCombined/corncob-output/Ex3 Affinity/all-responders.csv')
df.LI.norm$burn.trtmt <- factor(df.LI.norm$burn.trtmt, levels = c('control','wet','dry'))

for (i in 1:nrow(df.full3)) {
  if (df.full3$Estimate[i] > 0) {
    df.full3$Response[i] = 'positive'
  } else if (df.full3$Estimate[i] < 0) {
    df.full3$Response[i] = 'negative'
  }
}

df.full3.LIn <-  subset(df.full3, Comparison %in% c('LIn.dry.v.LIn.control'))$OTU
df.full3.LIwA <- subset(df.full3, Comparison %in% c('LIwA.dry.v.LIwA.control'))$OTU

df.count <- df.full3 %>%
  group_by(Comparison, Controlling.for) %>%
  count(Response)

# PUll out a subset of the corncob analyses:
df.affinity <- subset(df.full3, Comparison %in% c('LIwA.dry.v.LIwA.control') &
                        Controlling.for == 'pH category: low, mid, high' ) %>%
  subset(select = c(OTU, horizon, Response))

# OR just pull out positive responding OTUs:
df.affinity <- df.full3 %>%
  subset(Estimate > 0) %>%
  subset(Comparison %in% c('LIwA.dry.v.LIwA.control') & Controlling.for =='control pH') %>%
  subset(select = c(OTU, horizon, Response))

# Create a list of positive/negative responding OTUs
df.affinity.positive = subset(df.affinity, Response == 'positive')
df.affinity.negative = subset(df.affinity, Response == 'negative')

# Match these lists to the full community data.
df.LI.positive.responders <- subset(df.LI.norm, OTU %in% df.affinity.positive$OTU)
df.LI.negative.responders <- subset(df.LI.norm, OTU %in% df.affinity.negative$OTU)
df.LI.not.responders <- subset(df.LI.norm, !OTU %in% df.affinity$OTU)

# Create response categories to use in ggplot
df.LI.positive.responders$Response = 'positive'
df.LI.negative.responders$Response = 'negative'
df.LI.not.responders$Response = 'none'

# Combine the dataframe
df.LI.responders <- rbind(df.LI.positive.responders,df.LI.negative.responders,df.LI.not.responders)

# Convert pH to a character for the purpose of plotting
df.LI.responders$pH <- as.character(df.LI.responders$pH)
df.LI.responders$cntl.pH <- as.character(df.LI.responders$cntl.pH)

p1 = ggplot(subset(df.LI.responders, horizon == 'O' & site != 19), aes(x=cntl.pH, y=Abundance, fill=Response))+
  geom_bar(stat = 'identity') +
  facet_grid(incub.trtmt~burn.trtmt, scales = 'free', 
             labeller = labeller(incub.trtmt = IncubLabels, 
                                 horizon = HorizonLabels,
                                 burn.trtmt = BurnLabels)) +
  scale_fill_manual(values = c('forestgreen','grey50','red'),
                     limits = c('positive','none', 'negative'),
                     labels = c('Positive','No Response', 'Negative')) +
  labs(title = 'experiment3: Both LIn and LIwA data: controlling for pre-burn pH',
       y = 'Relative abundance') + 
  theme(legend.position = 'bottom',
        axis.text.x = element_text(size=8), 
        legend.text = element_text(size=8))

df.LI.sum <- df.LI.responders %>%
  group_by(Full.id, Response) %>%
  mutate(sum.abun = sum(Abundance)) %>%
  subset(select = c(Full.id, pH, cntl.pH, horizon, burn.trtmt, incub.trtmt, Response,site, sum.abun)) %>%
  unique()

df.LI.sum$burn.trtmt <- factor(df.LI.sum$burn.trtmt, levels = c('control','wet','dry'))
df.LI.sum$site <- as.character(df.LI.sum$site)

p2 = ggplot(subset(df.LI.sum, horizon == 'O' & Response != 'none' & burn.trtmt == 'dry'), aes(x=pH, y=(sum.abun), color = site))+
  geom_point(size=3) +
  facet_grid(incub.trtmt~burn.trtmt, scales = 'free', 
             labeller = labeller(incub.trtmt = IncubLabels, 
                                 horizon = HorizonLabels,
                                 burn.trtmt = BurnLabels)) +
        stat_smooth(method=lm,se = FALSE,color='black') +
  #scale_color_manual(values = c('forestgreen','grey50','red'),
   #                  limits = c('positive','none', 'negative'),
    #                 labels = c('Positive','No Response', 'Negative')) +
  labs(title = 'experiment3: Responders identified LIwA.dry.vs.control \ncontrolling for pre-burn pH',
       y = 'Relative abundance',
       x = 'post-burn pH') + 
  theme(legend.position = 'bottom',
        axis.text.x = element_text(size=8), 
        legend.text = element_text(size=8))+
  ylim(0,0.2)

```

# Mapping relative abundance of responding taxa across experiments:
```{r}
df.full.OTU <- read.csv('../data/sequence-data/LibCombined/corncob-output/all-resopnders.csv')

df.pb.gDNA.OTU <- read.csv('../data/sequence-data/LibCombined/phyloseq-objects/df.pb.gDNA.norm.csv')
df.SI.OTU <- read.csv('../data/sequence-data/LibCombined/phyloseq-objects/df.SI.norm.csv')
df.LI.norm <- read.csv('../data/sequence-data/LibCombined/phyloseq-objects/df.LI.norm.csv')

p = ggplot(data = df.pb.gDNA.OTU, aes(x=site, y = Abundance, fill = Phylum)) + 
  geom_bar(stat = 'identity') + 
  facet_grid(burn.trtmt~horizon) + 
  theme_bw()

df <- df.full.OTU %>%
  subset(Estimate > 0) %>%
  group_by(OTU) %>%
  mutate(max.Estimate = max(Estimate)) %>%
  subset(abs(Estimate) > 1) %>%
  merge(df.LI.norm) %>%
  subset(incub.trtmt == 'LIwA')

df$site <- as.character(df$site)
df$change.pH <- as.character(df$change.pH)
df$burn.trtmt <- factor(df$burn.trtmt, levels = c('control','wet','dry'))

p <- ggplot(data = subset(df, burn.trtmt == 'dry' & horizon == 'O'), 
            aes(x=site, y = Abundance, Fill = experiment)) + 
  theme_bw()+
  #geom_point()+
  geom_bar(stat = 'identity') +
  facet_wrap(~experiment, scale = 'free') + 
  #stat_smooth(method=lm,se = FALSE,color='black') +
  theme(axis.text.x = element_text(angle = 90))
```

# Exploring composition data for 24-hours post-burn:
```{r}
#df.pb.gDNA.OTU <- read.csv('../data/sequence-data/LibCombined/phyloseq-objects/df.pb.gDNA.norm.csv')
#df.pb.cDNA.OTU <- read.csv('../data/sequence-data/LibCombined/phyloseq-objects/df.pb.cDNA.norm.csv')

p = ggplot(data = df.pb.gDNA.OTU, aes(x=site, y = Abundance, fill = Phylum)) + 
  geom_bar(stat='identity')+
  facet_grid(horizon~burn.trtmt, labeller = labeller(horizon = HorizonLabels,
                                                     burn.trtmt = BurnLabels)) + 
  theme_bw() +
  theme(legend.position = 'bottom')

# Most abundant taxa; most abundant responders:
df <- df.pb.gDNA.OTU %>%
  group_by(Family) %>%
  mutate(sum.Abundance = sum(Abundance))

p = ggplot(data = subset(df, sum.Abundance > 2.08), aes(x=site, y = Abundance, fill = Class)) + 
  geom_bar(stat='identity')+
  facet_grid(horizon~burn.trtmt, labeller = labeller(horizon = HorizonLabels,
                                                     burn.trtmt = BurnLabels)) + 
  theme_bw() +
  theme(legend.position = 'bottom')

```

# Calculate bray-curtis dissimilarity for LIwA vs. SI
```{r, message = FALSE}
# This is a possibly convoluted piece of code to first create a Bray-Curtis
#     dissimilarity matrix and then pull out the dissimilarities between SI and
#     LIwA cores.
ps.bray <- prune_samples(sample_data(ps.raw.gDNA)$DNA.type == 'gDNA' &
                           sample_data(ps.raw.gDNA)$incub.trtmt %in% c('SI','LIwA')&
                           sample_data(ps.raw.gDNA)$Full.id %in% df.meta$Full.id, ps.raw.gDNA)

ps.bray <- prune_taxa(taxa_sums(ps.bray) > 0, ps.bray)

dist.bray.SIvLIwA <- as.matrix(distance(ps.bray, method = 'bray', type = 'samples'))

# Convert distance matrix into dataframe
df <- data.frame(dist.bray.SIvLIwA)

# Bring in meta data and subset for specific analysis
df.meta.sub <- subset(df.meta, select = c(Full.id, burn.trtmt,site, core.id.hor,
                                          incub.trtmt,core, horizon)) %>%
  subset(incub.trtmt %in% c('SI','LIwA')) %>%
  unique()

# Combine metadata and distances so we can use burn treatment to sort
#df <- merge(df, df.meta.pb.cDNA)

# Create a dataframe with the colnames and Full.ID
x <- data.frame(str_split(colnames(df)[1:276], 'X', simplify = TRUE)[,2])

# Rename columes:
colnames(x)[1] <- 'ID.bray'

# Clean up x. 
x <- x %>%
  mutate(Full.id = ID.bray)

x$Full.id <- str_replace_all(x$Full.id, "\\.", "-")

x$X <- 'X'

x <- x %>%
  unite(Bray.ID, c('X',"ID.bray"), sep = '', remove = TRUE) %>%
  unique()

# Create row with LIwA id, SI id, and core.id.hor:
df.LIwA <- subset(df.meta, incub.trtmt == 'LIwA') %>%
  mutate(LIwA.full.id = Full.id) %>%
  subset(select = c(core.id.hor, LIwA.full.id))

df.SI <- subset(df.meta, incub.trtmt == 'SI') %>%
  mutate(SI.full.id = Full.id) %>%
  subset(select = c(core.id.hor, SI.full.id)) %>%
  unique()

#df.list <- merge(df.LIwA, df.SI, by = 'core.id.hor')

# ANd then combine the Full.id list with the control core meta data I just created.
#     df.list now contains Full.id with the corresponding control sample for each 
#     dry/wet burn and a column ("Bray.ID") that has the sample ID formatted to 
#     match the distance matrix names.
df.list <- df.SI %>%
  merge(df.meta.sub, by = 'core.id.hor') %>%
  merge(x, by = 'Full.id') %>%
  subset(incub.trtmt == 'LIwA') %>%
  unique()

df.list$burn.trtmt <- as.character(df.list$burn.trtmt)
df.list$horizon <- as.character(df.list$horizon)
df.list$incub.trtmt <- as.character(df.list$incub.trtmt)

head(df.meta.sub)
# Create and empty data frame to store distances
df.distances <- data.frame("Full.id" = df.list$Full.id, 'core.id.hor' = 'core.id.hor',
                           'burn.trtmt' = 'burn', 'SI.full.id' = 'SI.full.id',
                           'horizon' = 'H',
                           'BC.distance.btw.SI.and.LIwA' = 'distance',
                           'incub.trtmt' = 'incub') %>%
  unique()

# Now run a convoluted loop:
for (i in 1:nrow(df.list)) { # For each sample (row in df.list)
  for (j in 1:nrow(df)) { # For each row in the dissimilarity matrix
    for (k in 1:ncol(df)){ # And for each colume in the dissimilarity matrix
      if(colnames(df)[k] == df.list$Bray.ID[i] &  # If the distance matrix colume ID matches the sample ID in the list
         rownames(df)[j] == df.list$SI.full.id[i]) { # And the rowname ID matches the corresponding control core, then...
        df.distances$BC.distance.btw.SI.and.LIwA[i] = df[j,k] # Pull out the distance and store it in new dataframe
        df.distances$burn.trtmt[i] = df.list$burn.trtmt[i]
        df.distances$horizon[i] = df.list$horizon[i]
        df.distances$incub.trtmt[i] = df.list$incub.trtmt[i]
        df.distances$SI.full.id[i] = df.list$SI.full.id[i]
      }
    } 
  } 
}

# Remove any samples that didn't have duplicates between LIwA and SI
df.distances <- subset(df.distances, horizon != 'H')

df.distances$BC.distance.btw.SI.and.LIwA <- as.numeric(df.distances$BC.distance.btw.SI.and.LIwA)

df.distances$incub.trtmt <- factor(df.distances$incub.trtmt, levels = c('pb','SI','LIwA','LIn'))

p=ggplot(subset(df.distances,horizon=='O')) + 
  geom_boxplot(aes(x=burn.trtmt, y = BC.distance.btw.SI.and.LIwA, fill = horizon),
               alpha = 0.5) +
  theme_bw() +
  facet_wrap(~horizon, ncol=3,
             labeller = labeller(horizon = HorizonLabels)) + 
  labs(x='Burn treatment',
       #y = 'Bray-Curtis dissimilarity between \nsoil 5 weeks post-burn vs. inoculated soil 6 months post-burn',
       y='Bray-Curtis dissimilarity',
       fill = 'Soil horizon',
       title = 'Bray-Curtis dissimilarities for gDNA') + 
  scale_fill_manual(values = c('black','grey80'),
                     limits = c('O','A'),
                     labels = c('Organic','Mineral')) +
  scale_x_discrete(limits = c('control','wet','dry'),
                   labels = c('Unburned \ncontrol', 'Wet soil \nburn','Dry soil \nburn')) +
  theme(axis.title = element_text(size=18),
        axis.text = element_text(size=18),
        legend.title = element_text(size=18),
        legend.text = element_text(size=18),
        strip.text = element_text(size=18))+
  ylim(0,1)

test <- aov(BC.distance.btw.SI.and.LIwA~burn.trtmt, data = subset(df.distances,horizon=='O'))
summary(test)
TukeyHSD(test)

df.meta.combined <- merge(df.meta, df.distances, by = c('Full.id', 'burn.trtmt', 'horizon', 'incub.trtmt'))

p = ggplot(subset(df.meta.combined, horizon == 'A'),aes(x=Thermo.mid.max, y = BC.distance.btw.burned.and.control, fill = horizon)) +
  geom_point(shape='circle filled',
             stroke=1,
             alpha = 0.8,
             size=3) +
  theme_bw() +
  facet_wrap(burn.trtmt~incub.trtmt, scales = 'free',
             labeller = labeller(incub.trtmt = IncubLabels,
                                 burn.trtmt = BurnLabels)) + 
  labs(x='Max temp, mid-point',
       y = 'Bray-Curtis dissimilarity compared to control; A horizon', 
       fill = 'Soil horizon',
       title = 'Bray-Curtis dissimilarities for gDNA') + 
  scale_fill_manual(values = c('black','grey40'),
                     limits = c('O','A'),
                     labels = c('Organic','Mineral')) +
  theme(axis.title = element_text(size=12),
        axis.text = element_text(size=12),
        legend.title = element_text(size=12),
        legend.text = element_text(size=12)) + 
    stat_smooth(method=lm,se = FALSE,color='black')

```

# Control pH vs. Delta pH:
```{r}
pChange_pH <- ggplot(subset(df.meta, burn.trtmt != 'control'),aes(x=cntl.pH, y = change.pH, color = Thermo.mid.max)) + 
  geom_point(aes(x=cntl.pH, y = change.pH, color = Thermo.mid.max), alpha = 0.4,size=3)  + 
      stat_smooth(method=lm,se = FALSE,color='black')+ 
    theme_bw()+
  labs(x= 'pre-burn (control) pH',
     #  y = 'Decay rate constant for fast C pool \nlog(k1)',
       y = 'Change in post-burn pH') + 
  facet_grid(burn.trtmt~horizon, scales = 'free_y',
             labeller = labeller(horizon = HorizonLabels,
                                                     burn.trtmt = BurnLabels))+
  #scale_color_manual(values = c('black','slategray','tan3'),
   #                  limits = c('control','wet','dry'),
    #                 labels = c('control','wet soil','dry soil')) +
  theme(axis.text.y = element_text(size = 14),
        legend.title = element_text(size=14),
        axis.text.x = element_text(face="italic",size = 14),
        axis.title.x = element_text(size = 14, margin = margin(t=10, b=10)),
        axis.title.y = element_text(size = 14, margin = margin(r=10, l=10)),
        strip.text = element_text(size=14),
        legend.text = element_text(size=14),
        legend.position = 'right',
       # legend.background = element_rect(fill = "darkgray"),
        legend.box = "vertical")

pChange_pH

df.meta.dry.O <- subset(df.meta, burn.trtmt == 'dry' & horizon == 'A')

model <- lm(change.pH~cntl.pH, data = df.meta.dry.O)
model
summary(model)

```

# Scatterplot matrices to explore "severity"
```{r}
colnames(df.meta)

my_colors <- c('red','orange','black')

df1 <- df.meta %>%
  subset(DNA.type =='gDNA' & incub.trtmt == 'pb') %>%
  mutate(Frac.O.loss = (PRE.O.hor.thickness.cm - POST.O.hor.thickness.cm)/PRE.O.hor.thickness.cm) %>%
  subset(select = c(burn.trtmt, 
                    PRE.O.hor.thickness.cm,
                    avg.sand,
                    Thermo.mid.max,
                    Thermo.low.max, 
                    change.pH, 
                    change.C.N.ratio,
                    Mid.degree.C.hours,
                    Low.degree.C.hours,
                    Frac.core.dry.mass.loss,
                    Frac.core.wet.mass.loss,
                    hrzn.degree.hrs,
                    Frac.O.loss))

pairs.panels(df1[,c(2:3,8:10)], 
      method = 'pearson',
      hist.col = 'blue',
      density = TRUE,
      ellipses = TRUE,
      col = my_colors[df1$burn.trtmt],
      pch=19)

pairs.panels(df1[,c(10,11,8,9)], 
      method = 'pearson',
      hist.col = 'blue',
      density = TRUE,
      ellipses = TRUE,
      col = my_colors[df1$burn.trtmt],
      pch=19)

pairs(df1[,c(2,3,4,6,7,8,12,13)],
      method = 'pearson',
      col = my_colors[df1$burn.trtmt],
      pch=19)

df.meta <- df.meta %>%
  mutate(Frac.O.loss = (PRE.O.hor.thickness.cm - POST.O.hor.thickness.cm)/PRE.O.hor.thickness.cm)

p3 = ggplot(df.meta, aes(x=hrzn.degree.hrs,y=Frac.O.loss, color=burn.trtmt)) +
  theme_bw()+
  geom_point(size=3) + 
  stat_smooth(method=lm,color='black') +
  facet_grid(~Veg.type)

```

# Comparing autoclave v. not decay model data:
```{r}
df.coefs <- subset(df.meta, incub.trtmt != 'SI' &
                     incub.trtmt != 'pb')

p1 = ggplot(df.coefs, aes(x=burn.trtmt)) + 
  geom_boxplot(aes(y=(M1), fill=incub.trtmt, color=incub.trtmt), alpha = 0.5, varwidth = FALSE)+
  #geom_boxplot(aes(y=log(M2), fill=burn.trtmt))+
  theme_bw()+
  labs(x= 'Dominant vegetation',
     #  y = 'Decay rate constant for fast C pool \nlog(k1)',
       y = 'Fractional size of fast C pool \n(M1)',
       fill = 'Burn treatment',
       color = 'Burn treatment') + 
  facet_wrap(~Veg.type) +
  theme(axis.text.y = element_text(size = 16),
        legend.title = element_text(size=16),
        axis.text.x = element_text(size = 16),
        axis.title.x = element_text(size = 16, margin = margin(t=10, b=10)),
        axis.title.y = element_text(size = 16, margin = margin(r=10, l=10)),
        strip.text = element_text(size=16),
        legend.text = element_text(size=16),
       # legend.background = element_rect(fill = "darkgray"),
        legend.box = "vertical")
p1

df.rsquared.5weeks <- read.csv('../data/2-pool-decay-model-output-5-weeks.csv')
df.5weeks <- subset(df.rsquared.5weeks, t == 3)

df.5weeks$incub.trtmt <- factor(df.5weeks$incub.trtmt, levels = c('SI','LIwA','LIn'))
df.5weeks$burn.trtmt <- factor(df.5weeks$burn.trtmt, levels = c('control','wet','dry'))

p1 = ggplot(df.5weeks, aes(x=burn.trtmt)) + 
  geom_boxplot(aes(y=(M1), fill=incub.trtmt, color=incub.trtmt), alpha = 0.5, varwidth = FALSE)+
  #geom_boxplot(aes(y=log(M2), fill=burn.trtmt))+
  theme_bw()+
  labs(y = 'Fractional size of fast C pool \n(M1)',
       fill = 'Burn treatment',
       color = 'Burn treatment') + 
  theme(axis.text.y = element_text(size = 16),
        legend.title = element_text(size=16),
        axis.text.x = element_text(size = 16),
        axis.title.x = element_text(size = 16, margin = margin(t=10, b=10)),
        axis.title.y = element_text(size = 16, margin = margin(r=10, l=10)),
        strip.text = element_text(size=16),
        legend.text = element_text(size=16),
       # legend.background = element_rect(fill = "darkgray"),
        legend.box = "vertical")
p1

# Look at fractional C remaining at end of incubation:
df.fracC <- read.csv("../data/fractional-c-remaining.csv", row.names = 1)

df.veg <- subset(df.meta, select = c(site, Veg.type)) %>%
  unique()

df.fracC <- subset(df.fracC, DayFactor == 34) %>%
  merge(df.veg, all = TRUE)

df.fracC$incub.trtmt <- factor(df.fracC$incub.trtmt, levels = c('SI','LIwA','LIn'))
df.fracC$burn.trtmt <- factor(df.fracC$burn.trtmt, levels = c('control','wet','dry'))

p = ggplot(df.fracC, aes(x = burn.trtmt)) + 
    geom_boxplot(aes(y=fractional.C.remaining, fill=incub.trtmt, color=incub.trtmt), alpha = 0.5, varwidth = FALSE)+
  theme_bw()+
  facet_wrap(~Veg.type)+
  labs(fill = 'Burn treatment',
       color = 'Burn treatment') + 
  theme(axis.text.y = element_text(size = 16),
        legend.title = element_text(size=16),
        axis.text.x = element_text(size = 16),
        axis.title.x = element_text(size = 16, margin = margin(t=10, b=10)),
        axis.title.y = element_text(size = 16, margin = margin(r=10, l=10)),
        strip.text = element_text(size=16),
        legend.text = element_text(size=16),
       # legend.background = element_rect(fill = "darkgray"),
        legend.box = "vertical")

```

# ORDINATIONS - Comparing to unburned
```{r}
MyNMDS.pb.gDNA <- ordinate(ps.norm.pb.gDNA, method = 'NMDS', k=3, ddistance="unifrac", weighted=TRUE)
MyNMDS.pb.cDNA <- ordinate(ps.norm.pb.cDNA, method = 'NMDS', k=3, ddistance="unifrac", weighted=TRUE)
MyNMDS.SI <- ordinate(ps.norm.SI, method = 'NMDS', k=3, ddistance="unifrac", weighted=TRUE)

MyNMDS.LI <- ordinate(ps.norm.LI, method = 'NMDS', k=3, ddistance="unifrac", weighted=TRUE)
MyNMDS.LI

ps.norm.LI.control <- prune_samples(sample_data(ps.norm.LI)$burn.trtmt %in% c('control'),ps.norm.LI)
ps.norm.LI.control<- prune_taxa(taxa_sums(ps.norm.LI.control)>0, ps.norm.LI.control)

ps.norm.LI.wet <- prune_samples(sample_data(ps.norm.LI)$burn.trtmt %in% c('wet'),ps.norm.LI)
ps.norm.LI.wet<- prune_taxa(taxa_sums(ps.norm.LI.wet)>0, ps.norm.LI.wet)

ps.norm.LI.dry <- prune_samples(sample_data(ps.norm.LI)$burn.trtmt %in% c('dry'),ps.norm.LI)
ps.norm.LI.dry<- prune_taxa(taxa_sums(ps.norm.LI.dry)>0, ps.norm.LI.dry)

MyNMDS.LI.control <- ordinate(ps.norm.LI.control, method = 'NMDS', k=3, ddistance="unifrac", weighted=TRUE)
MyNMDS.LI.wet <- ordinate(ps.norm.LI.wet, method = 'NMDS', k=3, ddistance="unifrac", weighted=TRUE)
MyNMDS.LI.dry <- ordinate(ps.norm.LI.dry, method = 'NMDS', k=3, ddistance="unifrac", weighted=TRUE)

MyNMDS.LI.control
MyNMDS.LI.wet
MyNMDS.LI.dry

p.LI.control = plot_ordination(ps.norm.LI.control, MyNMDS.LI.control,
                               axes=c(1,3),shape = "incub.trtmt", color = 'horizon') +
  geom_line(aes(group = core.id.hor)) +
  theme_bw() +
  geom_point(size=3)

p.LI.wet = plot_ordination(ps.norm.LI.wet, MyNMDS.LI.wet, 
                           axes = c(1,3),shape = "incub.trtmt", color = 'horizon') +  
  geom_line(aes(group = core.id.hor)) +

  theme_bw() +
  geom_point(size=3)

p.LI.dry = plot_ordination(ps.norm.LI.dry, MyNMDS.LI.dry, 
                           axes=c(1,3),shape = "incub.trtmt", color = 'horizon') +
    geom_line(aes(group = core.id.hor)) +
  theme_bw() +
  geom_point(size=3)

cowplot::plot_grid(p.LI.control,p.LI.wet,p.LI.dry, ncol=3, labels = c('Control','Wet','Dry'))

p.LI.axis1.2 = plot_ordination(ps.norm.LI, MyNMDS.LI, 
                           axes=c(1,2),shape = "horizon", color = 'burn.trtmt') +
    geom_line(aes(group = core.id.hor)) +
  theme_bw() +
  geom_point(size=3)
p.LI.axis1.3 = plot_ordination(ps.norm.LI, MyNMDS.LI, 
                           axes=c(1,3),shape = "horizon", color = 'burn.trtmt') +
    geom_line(aes(group = core.id.hor)) +
  theme_bw() +
  geom_point(size=3)
cowplot::plot_grid(p.LI.axis1.2,p.LI.axis1.3, ncol=2)

p1 = plot_ordination(ps.norm.pb.gDNA, MyNMDS.pb.gDNA, axes = c(1,2),shape = "burn.trtmt", color = 'burn.trtmt') +
  facet_wrap(~incub.trtmt) + theme_bw() +
  geom_point(size=3)
p2 = plot_ordination(ps.norm.pb.gDNA, MyNMDS.pb.gDNA, axes=c(1,3),shape = "burn.trtmt", color = 'burn.trtmt') +
  facet_wrap(~incub.trtmt) + theme_bw() +
  geom_point(size=3)
p3 = plot_ordination(ps.norm.SI, MyNMDS.SI, axes = c(1,2),shape = "burn.trtmt", color = 'burn.trtmt') +
  facet_wrap(~incub.trtmt) + theme_bw() +
  geom_point(size=3)
p4 = plot_ordination(ps.norm.SI, MyNMDS.SI, axes=c(1,3),shape = "burn.trtmt", color = 'burn.trtmt') +
  facet_wrap(~incub.trtmt) + theme_bw() +
  geom_point(size=3)
p5 = plot_ordination(ps.norm.LIwA, MyNMDS.LIwA, axes = c(1,2),shape = "burn.trtmt", color = 'burn.trtmt') +
  facet_wrap(~incub.trtmt) + theme_bw() +
  geom_point(size=3)
p6 = plot_ordination(ps.norm.LIwA, MyNMDS.LIwA, axes=c(1,3),shape = "burn.trtmt", color = 'burn.trtmt') +
  facet_wrap(~incub.trtmt) + theme_bw() +
  geom_point(size=3)

# using cowplot to combine multiple figures.
plot_grid(p1,p2,p3,p4,p5,p6,ncol=2)

# Full ordination
MyNMDS.full <- ordinate(ps.norm.gDNA, method = 'NMDS', k=3, ddistance="wunifrac", weighted=TRUE)

ps.norm.pb <- prune_samples(sample_data(ps.norm.full)$incub.trtmt == 'pb', ps.norm.full)
ps.norm.pb <- prune_taxa(taxa_sums(ps.norm.pb)>0, ps.norm.pb)
MyNMDS.pb <- ordinate(ps.norm.pb, method = 'NMDS', k=3, ddistance="wunifrac", weighted=TRUE)

sample_data(ps.norm.pb)$horizon = factor(sample_data(ps.norm.pb)$horizon, levels = c('O','A'))

# Ordination plot with horizons and deg.hrs
p1 = plot_ordination(ps.norm.pb, MyNMDS.pb, axes = c(1,2), 
                      color = "burn.trtmt",
                      shape = 'burn.trtmt') + 
  facet_grid(horizon~DNA.type, labeller = labeller(horizon = HorizonLabels,
                                                   DNA.type = DNALabels,
                                                   incub.trtmt = IncubLabels))+
  #facet_grid(~horizon, labeller = labeller(horizon = HorizonLabels))+
  geom_point(size=3, alpha=0.8) +
  labs(color = 'Burn treatment',
       #color = expression("Degree ("*degree*C*") hours"), 
       shape = 'Burn treatment') +
  #scale_color_gradientn(colours = c('gold2','red','red4','black')) +
  scale_color_manual(values = c('black','orange','red'),
                     labels = c('control', 'wet soil burn','dry soil burn'),
                     breaks = c('control','wet','dry'))+
  scale_shape_manual(values = c('circle','square','triangle'),
                     labels = c('control', 'wet soil burn','dry soil burn'),
                     breaks = c('control','wet','dry')) +
  #scale_shape_discrete(labels = c('Picea' = 'Picea spp.', 'Populus_tremuloides' = 'Populus tremuloides','Pinus_banksiana' = 'Pinus banksiana')) +
  #scale_shape_manual(values = c('circle','triangle'),
   #                  labels = c('O horizon','Mineral'),
    #                 breaks = c('O','A')) + 
  theme(legend.text = element_text(face = 'italic',size = 16),
        axis.title = element_text(size=16),
        axis.text = element_text(size=16),
        legend.title = element_text(size=16)) + 
  theme_bw()

p1

p2 = plot_ordination(ps.norm.gDNA, MyNMDS.full, axes = c(2,3), 
                      color = "pH",
                      shape = 'burn.trtmt') + 
  facet_grid(horizon~incub.trtmt, labeller = labeller(horizon = HorizonLabels))+
  #facet_grid(~horizon, labeller = labeller(horizon = HorizonLabels))+
  geom_point(size=3, alpha=0.8) +
  labs(color = 'post-burn pH',
       #color = expression("Degree ("*degree*C*") hours"), 
       shape = 'Burn treatment',
       title = "Ch 4, p2: NMDS ordination on weighted unifrac dissimilarities \n All DNA samples; all timepoints") +
  #scale_color_gradientn(colours = c('gold2','red','red4','black')) +
  #scale_shape_manual(values = c('circle','square','triangle'),
   #                  labels = c('control', 'wet soil burn','dry soil burn'),
    #                 breaks = c('control','wet','dry')) +
  #scale_shape_discrete(labels = c('Picea' = 'Picea spp.', 'Populus_tremuloides' = 'Populus tremuloides','Pinus_banksiana' = 'Pinus banksiana')) +
  #scale_shape_manual(values = c('circle','triangle'),
   #                  labels = c('O horizon','Mineral'),
    #                 breaks = c('O','A')) + 
  theme(legend.text = element_text(face = 'italic',size = 16),
        axis.title = element_text(size=16),
        axis.text = element_text(size=16),
        legend.title = element_text(size=16)) + 
  theme_bw()

p2
```

# ORDINATIONS - Comparing SI to LIwA - inoculated burned soils look more like uninoculated burned 5-week samples
```{r}
ps.norm.SI.v.LIwA <- prune_samples(sample_data(ps.norm.full)$incub.trtmt %in% c('SI','LIwA'), ps.norm.full)

ps.subset <- prune_samples(sample_data(ps.norm.SI.v.LIwA)$horizon == 'O' &
                             sample_data(ps.norm.SI.v.LIwA)$burn.trtmt != 'wet',ps.norm.SI.v.LIwA)

MyNMDS.inoculation <- ordinate(ps.subset, method = 'NMDS', k=3, ddistance="unifrac", weighted=TRUE)

IncubLabels = c('5 weeks post-burn', 'Inoculated soils \n6 months post-burn')
names(IncubLabels) = c('SI', 'LIwA')

p.inoc = plot_ordination(ps.subset, MyNMDS.inoculation,
                               axes=c(1,2),shape = "burn.trtmt", color='burn.trtmt') +
  geom_point(size=2)+
  geom_line(aes(group = site), color = 'grey60') +
  facet_wrap(~incub.trtmt, labeller = labeller(incub.trtmt=IncubLabels))+
  scale_color_manual(values = c('black','tan3'),
                     limits = c('control','dry'),
                     labels = c('Unburned \ncontrol','Dry soil')) +
  scale_shape_manual(values = c('circle','triangle'),
                     limits = c('control','dry'),
                     labels = c('Unburned \ncontrol','Dry soil')) +
  labs(shape = 'Burn treatment',
       color = 'Burn treatment')+
  theme_bw() +
  geom_point(size=3)

p.inoc
```

# Phylogenetic tree:
```{r, message = FALSE}
ps.traits <- phyloseq::prune_taxa(df.trait$OTU, ps.raw.full)

AbunTaxa = taxa_names(filter_taxa(ps.norm.gDNA, function(x) mean(x) > 0.001, TRUE))

length(AbunTaxa)
ps.prune = prune_taxa(AbunTaxa, ps.raw.full)
ps.prune

ps.prune <- ps.raw.full %>%
  subset_samples(burn.trtmt %in% c('dry'))

ps.prune <- prune_taxa(AbunTaxa, ps.prune)

p = ggtree(ps.prune, aes(color=Phylum), 
           branch.length = 'none',
           layout = 'circular') + 
  geom_treescale() + 
  geom_tiplab2(size=2, aes(angle=angle))

tree <- read.tree(file = '../../CombinedFireSim1519/tree/merged-exported-tree/tree.nwk')
tree <- phy_tree(ps.prune)

plot(tree, type = 'cladogram')

# using cowplot to combine multiple figures.
cowplot::plot_grid(p,p1, ncol=2, labels = c('A','B'))

```

# Number of responding OTUs for each experiment:
```{r}
df.full.OTU <- read.csv('../data/sequence-data/LibCombined/corncob-output/all-responders.csv')
head(df.full.OTU)

# Filter responders:
df.full <- df.full.OTU %>%
  subset(abs(Estimate) > 1) # removes 1500 non-unique OTUs
dim(df.full)

df.PB.pos.rspd <- subset(df.full, experiment == 'survival' & Estimate > 0) %>%
  subset(Temp.cutoff %in% c('Low thermo greater than 50 C',
                            'Mid thermo greater than 50 C')) %>%
  group_by(OTU,horizon) %>%
  mutate(max.Estimate = max(Estimate)) %>%
  subset(max.Estimate == Estimate) %>%  
  mutate(Response = 'positive') %>%
  unique()

df.PB.neg.rspd <- subset(df.full, experiment == 'survival' & Estimate < 0) %>%
  subset(Temp.cutoff %in% c('Low thermo greater than 50 C',
                            'Mid thermo greater than 50 C')) %>%
  group_by(OTU, horizon) %>%
  mutate(min.Estimate = min(Estimate)) %>%
  subset(min.Estimate == Estimate) %>%
  mutate(Response = 'negative') %>%
  unique()

summary(df.PB.pos.rspd$Estimate)
length(unique(df.PB.pos.rspd$OTU))
length(unique(df.PB.neg.rspd$OTU))

df.SI.cntl.pos.rspd <- subset(df.full, experiment == 'fast growth' & Estimate > 0) %>%
  subset(Comparison == 'pb.control.v.SI.control') %>%
  group_by(OTU, horizon) %>%
  mutate(max.Estimate = max(Estimate)) %>%
  subset(max.Estimate == Estimate) %>%
  unique()
df.SI.cntl.neg.rspd <- subset(df.full, experiment == 'fast growth' & Estimate < 0) %>%
  subset(Comparison == 'pb.control.v.SI.control') %>%
  group_by(OTU, horizon) %>%
  mutate(min.Estimate = min(Estimate)) %>%
  subset(min.Estimate == Estimate) %>%
  unique()

df.SI.pos.rspd <- subset(df.full, experiment == 'fast growth' & Estimate > 0) %>%
  subset(Temp.cutoff %in% c('Low thermo greater than 50 C',
                            'Mid thermo greater than 50 C')) %>%
  group_by(OTU, horizon) %>%
  mutate(max.Estimate = max(Estimate)) %>%
  subset(max.Estimate == Estimate) %>%
  mutate(Response = 'positive') %>%
  unique() %>% 
  subset(!(OTU %in% df.SI.cntl.pos.rspd$OTU))
length(unique(df.SI.pos.rspd$OTU))

df.SI.neg.rspd <- subset(df.full, experiment == 'fast growth' & Estimate < 0) %>%
  subset(Temp.cutoff %in% c('Low thermo greater than 50 C',
                            'Mid thermo greater than 50 C')) %>%
  group_by(OTU, horizon) %>%
  mutate(min.Estimate = min(Estimate)) %>%
  subset(min.Estimate == Estimate) %>%
  unique() %>%
  mutate(Response = 'negative') %>%
  subset(!(OTU %in% df.SI.cntl.neg.rspd$OTU))

length(unique(df.SI.neg.rspd$OTU))

# LIwA:
df.LIwA <- subset(df.full, experiment == 'affinity')

df.LIwA.pos.rspd <- subset(df.full, experiment == 'affinity' & Estimate > 0) %>%
  subset(Temp.cutoff %in% c('Low thermo greater than 50 C',
                            'Mid thermo greater than 50 C')) %>%
  group_by(OTU, horizon) %>%
  mutate(max.Estimate = max(Estimate)) %>%
  subset(max.Estimate == Estimate) %>%
  mutate(Response = 'positive') %>%
  unique() 
length(unique(df.LIwA.pos.rspd$OTU))

df.LIwA.neg.rspd <- subset(df.full, experiment == 'affinity' & Estimate < 0) %>%
  subset(Temp.cutoff %in% c('Low thermo greater than 50 C',
                            'Mid thermo greater than 50 C')) %>%
  group_by(OTU,horizon) %>%
  mutate(min.Estimate = min(Estimate)) %>%
  subset(min.Estimate == Estimate) %>%
  mutate(Response = 'negative') %>%
  unique() 
length(unique(df.LIwA.neg.rspd$OTU))

df.responders <- rbind(df.PB.pos.rspd, df.PB.neg.rspd,
                       df.SI.pos.rspd, df.SI.neg.rspd,
                       df.LIwA.pos.rspd, df.LIwA.neg.rspd)
head(df.responders)

p = ggplot(subset(df.responders, Phylum == 'Firmicutes'), aes(x=Estimate, y = Genus, color = experiment)) + 
  geom_point(size=2) + 
  theme_bw() + 
  facet_grid(horizon~experiment) + 
  geom_vline(xintercept = 0)

# Are fire survivors likely to be fast growing?
length(df.PB.pos.rspd$OTU)
dim(subset(df.PB.pos.rspd, OTU %in% df.SI.pos.rspd$OTU))
# --> No not really. Of 19 fire survivors, only 2 are fast growing.

# Are fast growing taxa also survivors?
length(unique(df.SI.pos.rspd$OTU))
dim(subset(df.SI.pos.rspd, OTU %in% df.PB.pos.rspd$OTU)) 
# --> No, of 98 OTUs identified as fast growing, only 2 were also identified as fire surviving.

# Do fast growing taxa show post-fire affinity?
length(unique(df.SI.pos.rspd$OTU))
length(unique(subset(df.SI.pos.rspd, OTU %in% df.LIwA.pos.rspd$OTU))$OTU)
length(unique(subset(df.SI.pos.rspd, OTU %in% df.LIwA.neg.rspd$OTU))$OTU)
# --> 8 of 98 are positive+positive, 
# --> 5 of 98 are positive+negative

# Are positive affinity taxa also fire susceptible?
length(df.LIwA.pos.rspd$OTU)
dim(subset(df.LIwA.pos.rspd, OTU %in% df.PB.neg.rspd$OTU))
# --> Nope. Only 1 of the positive affinity taxa was identified as fire susceptible.

# Are positive affinity taxa also fast growers?
length(df.LIwA.pos.rspd$OTU)
dim(subset(df.LIwA.pos.rspd, OTU %in% df.SI.pos.rspd$OTU))
# --> Not really, 7 of 53 positive affinity taxa identified as fast growing.

# Are negative affinity taxa the same as slow growing taxa?
length(df.LIwA.neg.rspd$OTU) 
dim(subset(df.LIwA.neg.rspd, OTU %in% df.SI.neg.rspd$OTU))
# Not really, 12 of 199 (10%) of negative affinity taxa were identified as slow growing.

# DO fire survivors appear to have an affinity for post-fire environment?
length(df.PB.pos.rspd$OTU)
dim(subset(df.PB.pos.rspd, OTU %in% df.LIwA.pos.rspd$OTU))
# --> Nope. 1 of 19

# What about an anti-affinity?
dim(subset(df.PB.pos.rspd, OTU %in% df.LIwA.neg.rspd$OTU))
# --> Nop3. 2 of 19.
```

# Attempt at 3D plot
```{r}
# GRoup by Order
ps.norm.gDNA <- ps.norm.gDNA %>%
  tax_glom('Phylum')

df.norm.gDNA.Phylum <- psmelt(ps.norm.gDNA)

df.Phylum.pb <- subset(df.norm.gDNA.Phylum, incub.trtmt == 'pb') %>%
  group_by(Phylum, core.id.hor) %>%
  mutate(mean.Abun.pb = mean(Abundance)) %>%
  subset(select = c(Domain, Phylum, mean.Abun.pb, core.id.hor, burn.trtmt))

df.Phylum.SI <- subset(df.norm.gDNA.Phylum, incub.trtmt == 'SI') %>%
  group_by(Phylum, core.id.hor) %>%
  mutate(mean.Abun.SI = mean(Abundance)) %>%
  subset(select = c(Domain, Phylum, mean.Abun.SI, core.id.hor, burn.trtmt))

df.Phylum.LIn <- subset(df.norm.gDNA.Phylum, incub.trtmt == 'LIn') %>%
  group_by(Phylum, core.id.hor) %>%
  mutate(mean.Abun.LIn = mean(Abundance)) %>%
  subset(select = c(Domain, Phylum, mean.Abun.LIn, core.id.hor, burn.trtmt))

df.Phylum.LIwA <- subset(df.norm.gDNA.Phylum, incub.trtmt == 'LIwA') %>%
  group_by(Phylum, core.id.hor) %>%
  mutate(mean.Abun.LIwA = mean(Abundance)) %>%
  subset(select = c(Domain, Phylum, mean.Abun.LIwA, core.id.hor, burn.trtmt))

df <- merge(df.Phylum.pb, df.Phylum.SI) %>%
  merge(df.Phylum.LIn) %>%
  merge(df.Phylum.LIwA) %>%
  subset(mean.Abun.pb > 0.1 |
           mean.Abun.SI > 0.1 |
           mean.Abun.LIn > 0.1)

# Plot
p = ggplot(df, aes(x=mean.Abun.pb, 
                   y=mean.Abun.SI, 
                   color=Phylum, 
                   shape = burn.trtmt,
                   size = mean.Abun.LIn))+
  geom_point() + 
  #facet_wrap(~Phylum, scales = 'fixed') + 
  ylim(0,1) + 
  xlim(0,1)
p

cols <- c('red','orange','yellow','green','blue','purple','pink','black','grey10','grey20', 'forestgreen')

df$color <- mycolors[as.numeric(df$Phylum)]

# 3D plot with scatterplot3d - problems with coloring by variable
with(df,
     scatterplot3d(x = mean.Abun.pb, 
                  y = mean.Abun.SI,
                  z = mean.Abun.LIn))

# 3D plot with rgl
plot3d(
  x=df$mean.Abun.pb, y=df$mean.Abun.SI, z=df$mean.Abun.LIn,
  type = 's',
  raidus = 0.1)
```