### 2023-03-16 README.rmd
Update README file

Scripts for analyzing data associated with "Experimentally-determined traits shape bacterial community composition one and five years following wildfire"

corncob-experiment1.R: tests for differentially abundant OTUs in unburned soil vs. soil 24 hrs post-burn using the corncob package (Martin et al. 2020)

corncob-experiment2-TempCutoff.R: tests for differentially abundant OTUs in unburned/burned soils after a 24 hr vs. 5 week incubation using corncob

corncob-experiment3.R: tests for differentially abundant OTUs in burned vs. unburned soils following autoclaving, inoculation with unburned soil, and a 6 month incubation

create-phyloseq.R: takes seq data in form of BIOM table and list of sequence IDs and cleans up samples by removing blanks, lab standard, duplicated samples. Output is a phyloseq object with sample ID, taxonomic table, OTU table, and phylogenetic tree.

creating_burn_temp_profiles.R: plot the temperature profiles during simulated burns.

decay-model-two-pool.R: Fits two pool decay model to CO2 respiration data

incubations_microbial_respiration.R: calculate C remaining during laboratory incubations as a fraction of initial total C

Manuscript-figrues.Rmd: Script to create figures in the manuscript and SI

Merged_paper_figures_Revision_Github.R: Combine lab and field OTU datasets to run ordinations and quantify relative abundance of laboratory-identified trait taxa in field data

OTU-trait-list.R - compiles list of OTUs identified as differentially enriched from each experiment

Qiime-pipeline-for-FireSim2019-LibCombined.ipynb: steps for sequence processing and taxonomic assignments

rrndb-SI.R: Assigns mean predicted 16S copy numbers to OTUs based on taxonomy